# do_BarPlot() | Categorical proportions

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Bar plots display counts or proportions of categorical variables. Use them to compare cell type composition across samples, conditions, or any grouping variable.

---

## Basic usage

Count cells per cluster:

```{r}
#| eval: true
p <- SCpubr::do_BarPlot(sample = sample,
                        group.by = "Annotation")
p
```

---

## Split by a second variable

Show how clusters are distributed across samples:

```{r}
#| eval: true
p <- SCpubr::do_BarPlot(sample = sample,
                        group.by = "Annotation",
                        split.by = "ID",
                        position = "stack")
p
```

Think of it as:
- `group.by`: What creates the bars (total counts per value)
- `split.by`: How each bar is subdivided

---

## Show proportions

Use `position = "fill"` to normalize bars to the same height:

```{r}
#| eval: true
p <- SCpubr::do_BarPlot(sample = sample,
                        group.by = "Annotation",
                        split.by = "ID",
                        position = "fill",
                        flip = TRUE)
p
```

---

## Add total counts

Display the total cell count on each bar:

```{r}
#| eval: true
p <- SCpubr::do_BarPlot(sample = sample,
                        group.by = "Annotation",
                        split.by = "ID",
                        position = "fill",
                        flip = TRUE,
                        add.n = TRUE,
                        add.n.size = 4,
                        add.n.expand = c(0, 1.15))
p
```

---

## Order bars

Order by total count or by a specific group's value:

```{r}
#| eval: true
# Order by total count
p <- SCpubr::do_BarPlot(sample = sample,
                        group.by = "Annotation",
                        order = TRUE)

p
```

---

## Facet by a third variable

Break down bars by an additional grouping:

```{r}
#| eval: true
p <- SCpubr::do_BarPlot(sample = sample,
                        group.by = "Annotation",
                        split.by = "ID",
                        facet.by = "subtype",
                        position = "fill",
                        flip = TRUE)
p
```

---


## Return data

Get the underlying count/proportion data:

```{r}
#| eval: true
result <- SCpubr::do_BarPlot(sample = sample,
                             group.by = "Annotation",
                             split.by = "ID",
                             return_data = TRUE)

head(result$Data)
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `group.by` | Metadata column for bars | — |
| `split.by` | Metadata column to subdivide bars | `NULL` |
| `facet.by` | Metadata column to facet by | `NULL` |
| `position` | `"stack"` (counts) or `"fill"` (proportions) | `"stack"` |

### Ordering

| Parameter | Description | Default |
|-----------|-------------|---------|
| `order` | Order bars by descending count | `FALSE` |
| `order.by` | Value of `split.by` to order by | `NULL` |

### Count labels

| Parameter | Description | Default |
|-----------|-------------|---------|
| `add.n` | Show total counts on bars | `FALSE` |
| `add.n.size` | Label text size | `4` |
| `add.n.face` | Label font face | `"bold"` |
| `add.n.expand` | Y-axis expansion for labels | `c(0, 1.15)` |

### Output

| Parameter | Description | Default |
|-----------|-------------|---------|
| `return_data` | Return data instead of plot | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_WafflePlot()](02_waffleplot.qmd) — Pictogram proportions
- [do_AlluvialPlot()](03_alluvialplot.qmd) — Flow between categories
