# do_WafflePlot() | Pictogram proportions

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Waffle plots display proportions as a **10×10 grid** where each cell represents 1% of the data. This pictogram format makes it easy to intuitively grasp relative proportions.

---


---

## Basic usage

```{r}
#| eval: true
p <- SCpubr::do_WafflePlot(sample = sample,
                           group.by = "subtype")
p
```

Each tile in the 10×10 grid represents 1% of cells. The grid is filled proportionally by each cluster's frequency.

---

## Customize tile size

Adjust the spacing between tiles:

```{r}
#| eval: true
# Thicker borders
p <- SCpubr::do_WafflePlot(sample = sample,
                           group.by = "subtype",
                           waffle.size = 4)

p
```

```{r}
#| eval: true
# Thinner borders
p <- SCpubr::do_WafflePlot(sample = sample,
                           group.by = "subtype",
                           waffle.size = 1)
p
```

---

## Orientation

```{r}
#| eval: true
# Horizontal (default)
p <- SCpubr::do_WafflePlot(sample = sample,
                           group.by = "subtype",
                           flip = TRUE)

p
```

---

## Custom colors

```{r}
#| eval: true
colors <- c("ATRT-TYR" = "#E63946", "ATRT-SHH" = "#457B9D", "ATRT-MYC" = "#1D3557")

p <- SCpubr::do_WafflePlot(sample = sample,
                           group.by = "subtype",
                           colors.use = colors)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `group.by` | Metadata column to visualize | — |
| `waffle.size` | Border size between tiles | `2` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_BarPlot()](01_barplot.qmd) — Bar chart proportions
- [do_AlluvialPlot()](03_alluvialplot.qmd) — Flow between categories
