# do_AlluvialPlot() | Flow between categories

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Alluvial plots visualize **how cells flow between categorical groups**—for example, how cluster assignments change across conditions, time points, or annotation methods. Built on the `ggalluvial` package.

---

## Basic usage

Visualize how cells in clusters map to cell types:

```{r}
#| eval: true
p <- SCpubr::do_AlluvialPlot(sample = sample,
                             first_group = "Annotation",
                             last_group = "subtype")
p
```

---

## Multiple intermediate groups

Add intermediate nodes between start and end:

```{r}
#| eval: true
p <- SCpubr::do_AlluvialPlot(
  sample        = sample,
  first_group   = "technology",
  middle_groups = c("Annotation"),
  last_group    = "subtype"
)
p
```

---

## Color by group

Control which group determines the flow colors:

```{r}
#| eval: true
# Color by last group (default)
p <- SCpubr::do_AlluvialPlot(sample = sample,
                             first_group = "Annotation",
                             last_group = "subtype",
                             fill.by = "subtype")

p
```
```{r}
#| eval: true
# Color by first group 
p <- SCpubr::do_AlluvialPlot(sample = sample,
                             first_group = "Annotation",
                             last_group = "subtype",
                             fill.by = "Annotation")

p
```
---

## Stratum customization

### Fill stratum based on flow

```{r}
#| eval: true
p <- SCpubr::do_AlluvialPlot(sample = sample,
                             first_group = "Annotation",
                             last_group = "subtype",
                             stratum.fill.conditional = TRUE)
p
```

### Custom stratum appearance

```{r}
#| eval: true
p <- SCpubr::do_AlluvialPlot(sample = sample,
                             first_group = "Annotation",
                             last_group = "subtype",
                             stratum.fill = "grey90",
                             stratum.color = "black",
                             stratum.width = 1/4)
p
```

---

## Flow styles

### Use geom_flow (alternative rendering)

```{r}
#| eval: true
p <- SCpubr::do_AlluvialPlot(sample = sample,
                             first_group = "Annotation",
                             last_group = "subtype",
                             use_geom_flow = TRUE)
p
```

### Curve types

```{r}
#| eval: true
# Available: linear, cubic, quintic, sine, arctangent, sigmoid, xspline
p <- SCpubr::do_AlluvialPlot(sample = sample,
                             first_group = "Annotation",
                             last_group = "subtype",
                             curve_type = "xspline")
p
```

---

## Use labels

Replace stratum boxes with text labels:

```{r}
#| eval: true
p <- SCpubr::do_AlluvialPlot(sample = sample,
                             first_group = "Annotation",
                             last_group = "subtype",
                             use_labels = TRUE,
                             repel = TRUE)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `first_group` | Starting categorical variable | — |
| `last_group` | Ending categorical variable | — |
| `middle_groups` | Intermediate variables | `NULL` |
| `fill.by` | Variable to color flows by | `last_group` |

### Stratum appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `stratum.fill` | Stratum fill color | `"white"` |
| `stratum.color` | Stratum border color | `"black"` |
| `stratum.width` | Stratum width | `1/3` |
| `stratum.fill.conditional` | Match stratum to flow color | `FALSE` |

### Flow appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `use_geom_flow` | Use alternative flow rendering | `FALSE` |
| `alluvium.color` | Flow border color | `"white"` |
| `curve_type` | Curve shape | `"sigmoid"` |

### Labels

| Parameter | Description | Default |
|-----------|-------------|---------|
| `use_labels` | Show text labels instead of boxes | `FALSE` |
| `repel` | Repel overlapping labels | `FALSE` |
| `label.color` | Label text color | `"black"` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_BarPlot()](01_barplot.qmd) — Bar chart proportions
- [do_ChordDiagramPlot()](04_chorddiagramplot.qmd) — Circular flow visualization
