# do_ChordDiagramPlot() | Circular flow visualization

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Chord diagrams display **relationships between categorical groups** as curved links in a circular layout. Built on the `circlize` package, they're ideal for visualizing cell type transitions or interactions.

---

## Basic usage

Visualize transitions between two categorical variables:

```{r}
#| eval: true
p <- SCpubr::do_ChordDiagramPlot(sample = sample,
                                 from = "Annotation",
                                 to = "subtype")
p
```

---

## Directional links

Control link direction:

```{r}
#| eval: true
# From → To (default)
p <- SCpubr::do_ChordDiagramPlot(sample = sample,
                                 from = "Annotation",
                                 to = "subtype",
                                 directional = 1)

p
```

```{r}
#| eval: true
# To → From
p <- SCpubr::do_ChordDiagramPlot(sample = sample,
                                 from = "Annotation",
                                 to = "subtype",
                                 directional = -1)
p
```

---

## Direction visualization

```{r}
#| eval: true
# Height difference + arrows (default)
p <- SCpubr::do_ChordDiagramPlot(sample = sample,
                                 from = "Annotation",
                                 to = "subtype",
                                 direction.type = c("diffHeight", "arrows"))

p
```

```{r}
#| eval: true
# Arrow types: "triangle" or "big.arrow"
p <- SCpubr::do_ChordDiagramPlot(sample = sample,
                                 from = "Annotation",
                                 to = "subtype",
                                 link.arr.type = "triangle")
p
```

---

## Highlight a group

Focus attention on links from a specific group:

```{r}
#| eval: true
p <- SCpubr::do_ChordDiagramPlot(sample = sample,
                                 from = "Annotation",
                                 to = "subtype",
                                 highlight_group = "CP-like",
                                 alpha.highlight = 25)
p
```

---

## Scale nodes equally

Make all groups the same width regardless of cell count:

```{r}
#| eval: true
p <- SCpubr::do_ChordDiagramPlot(sample = sample,
                                 from = "Annotation",
                                 to = "subtype",
                                 scale = TRUE)
p
```


---

## Alignment

```{r}
#| eval: true
# Horizontal split between groups
p <- SCpubr::do_ChordDiagramPlot(sample = sample,
                                 from = "Annotation",
                                 to = "subtype",
                                 alignment = "horizontal")

p
```

```{r}
#| eval: true
# Vertical split
p <- SCpubr::do_ChordDiagramPlot(sample = sample,
                                 from = "Annotation",
                                 to = "subtype",
                                 alignment = "vertical")
p
```


---

## Parameter reference

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `from` | Source categorical variable | — |
| `to` | Target categorical variable | — |
| `colors.from` | Colors for source groups | Auto |
| `colors.to` | Colors for target groups | Auto |

### Directionality

| Parameter | Description | Default |
|-----------|-------------|---------|
| `directional` | Direction: 0 (none), 1 (from→to), -1 (to→from), 2 (both) | `1` |
| `direction.type` | How to show direction | `c("diffHeight", "arrows")` |
| `link.arr.type` | Arrow style: `"triangle"` or `"big.arrow"` | `"big.arrow"` |

### Layout

| Parameter | Description | Default |
|-----------|-------------|---------|
| `big.gap` | Gap between from/to groups | `10` |
| `small.gap` | Gap within groups | `1` |
| `scale` | Equal width for all groups | `FALSE` |
| `alignment` | `"default"`, `"horizontal"`, `"vertical"` | `"default"` |
| `self.link` | 1 (prevent) or 2 (allow) self-links | `1` |

### Highlighting

| Parameter | Description | Default |
|-----------|-------------|---------|
| `highlight_group` | Group to highlight | `NULL` |
| `alpha.highlight` | Highlighted link transparency (00-99) | `25` |
| `z_index` | Bring larger links to top | `FALSE` |

### Link appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `link.border.color` | Link border color | `NA` |
| `link.border.width` | Link border width | `1` |

---

## See also

- [do_AlluvialPlot()](03_alluvialplot.qmd) — Linear flow visualization
- [do_LigandReceptorPlot()](../05_downstream_analyses/09_ligandreceptorplot.qmd) — Ligand-receptor interactions
