# do_DimPlot() | Dimensional Reduction scatter plots

```{r}
#| results: "asis"
#| echo: false
source("_common.R")
suppressMessages(library("Seurat"))
```

```{r}
#| echo: false
sample <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/sc_dataset.rds")
```


Dimensional reduction Plots (`DimPlots`) are a highly recognizable visualization in single-cell experiments. They allow users to visualize cells in a dimensional reduction embedding, such as `PCA` or `UMAP`. Users can color cells according to any desired groups, enabling visualization of any kind of categorical data on the cells in the dimensional reduction embedding.

## Basic usage

DimPlots can be generated in `SCpubr` using the function `SCpubr::do_DimPlot()`:

```{r}
#| fig-width: 10
#| fig-height: 11

p <- SCpubr::do_DimPlot(sample = sample)

p
```

## Modifying axes behavior

Axes are removed by default from the plot, together with the axes titles. This behaviour can be overridden with `plot.axes = TRUE`:

```{r}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Bring back the axes."

# Bring back the Axes.
p <- SCpubr::do_DimPlot(sample = sample,
                        plot.axes = TRUE)
p
```


## Label the clusters

In some cases,  we might want to remove the legend entirely, and instead plot labels on top of each cluster. This can be achieved by using `label = TRUE`.

```{r}
#| fig-width: 10
#| fig-height: 10

# Put labels on top of the clusters.
p <- SCpubr::do_DimPlot(sample, 
                     label = TRUE)
p
```

These labels are, in essence the result of applying `ggplot2::geom_label()` on the plot. However, we might also want to have them as pure text instead of the labels. We can achieve that by providing `label.box = FALSE`.

```{r}
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Labels as text."

# Labels as text
p <- SCpubr::do_DimPlot(sample = sample, 
                        label = TRUE,
                        label.box = FALSE)
p
```

However, we can play further with other parameters of the function such as `label.color` which will provide a different color for the text inside the labels and `label.fill` to modify the background of the labels.:

```{r}
#| fig-width: 20
#| fig-height: 10

# Change the color of the label text.
p1 <- SCpubr::do_DimPlot(sample = sample, 
                         label = TRUE, 
                         label.color = "white",
                         label.fill = "black")

# Change the color of the text.
p2 <- SCpubr::do_DimPlot(sample = sample, 
                         label = TRUE, 
                         label.color = "black",
                         label.box = FALSE)
p <- p1 | p2
p
```

Also, we can modify the size of the labels/text by using `label.size` paramter:

```{r}
#| fig-width: 20
#| fig-height: 10

# Change the size of the label text.
p1 <- SCpubr::do_DimPlot(sample = sample, 
                         label = TRUE, 
                         label.size = 6)

# Change the size of the text.
p2 <- SCpubr::do_DimPlot(sample = sample, 
                         label = TRUE, 
                         label.box = FALSE,
                         label.size = 6)
p <- p1 | p2
p
```

Finally, if the labels/text are overlapping one to another, we can try to fix that by repelling the labels from one to another. We can achieve this by using `repel = TRUE`:

```{r}
#| fig-width: 20
#| fig-height: 10

# Repel the labels.
p1 <- SCpubr::do_DimPlot(sample = sample, 
                         label = TRUE, 
                         label.color = "black",
                         repel = TRUE)

# Repel the text.
p2 <- SCpubr::do_DimPlot(sample = sample, 
                         label = TRUE, 
                         label.color = "black",
                         label.box = FALSE,
                         repel = TRUE)
p <- p1 | p2
p
```


## Changing the order of plotting

Be default `SCpubr::do_DimPlot()` plots cells randomly using `shuffle = TRUE`. This is different from the default behavior of `Seurat::DimPlot()`, which plots cells based on their identity factor levels.

```{r}
#| fig-width: 20
#| fig-height: 11

p1 <- SCpubr::do_DimPlot(sample = sample,
                         reduction = "pca",
                         shuffle = TRUE)

p2 <- SCpubr::do_DimPlot(sample = sample,
                         reduction = "pca",
                         shuffle = FALSE)

p <- p1 | p2
p
```


## Highlighting cells

We can highlight a certain group of cells in the plot. This is achieved by using the `cells.highlight` parameter.

```{r}
#| fig-width: 10
#| fig-height: 11

cells.use <- sample(x = colnames(sample), 
                    size = 1500)

p <- SCpubr::do_DimPlot(sample = sample,
                        cells.highlight = cells.use)

p
```

One can also change the color of the highligted cells by providing single color to `colors.use` and the color of the not selected cells with `na.value`:

```{r}
#| fig-width: 10
#| fig-height: 11

# Change color of highlighted and non-highlighted cells.
p <- SCpubr::do_DimPlot(sample = sample, 
                        cells.highlight = cells.use,
                        colors.use = "dodgerblue",
                        na.value = "grey90")
p
```

The size of the highlighted dots can be modified with the parameter `sizes.highlight`.

```{r}
#| fig-width: 10
#| fig-height: 11

# Increase the size of the highlighted cells.
p <- SCpubr::do_DimPlot(sample = sample, 
                        cells.highlight = cells.use, 
                        sizes.highlight = 2)
p
```

We can also highlight **whole identities** with `idents.highlight` parameter. For this, just provide the desired identities to be selected. It can also work in combination with `cells.highlight`.

```{r}
#| fig-width: 30
#| fig-height: 11

# Using cells.highlight.
p1 <- SCpubr::do_DimPlot(sample = sample, 
                         cells.highlight = cells.use)

# Using idents.highlight.
p2 <- SCpubr::do_DimPlot(sample = sample, 
                         idents.highlight = c("6"))

# Using both.
p3 <- SCpubr::do_DimPlot(sample = sample, 
                         cells.highlight = cells.use, 
                         idents.highlight = c("6"))

p <- p1 | p2 | p3
p
```

## Restrict the identitites displayed

Sometimes, we are interested in showing only some of the identities or groups in our sample. Instead of highlighting cells, we still want to keep the original colors and legend. For this use case, one might be tempted to just subset the sample as follows:

```{r}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Subsetting the Seurat object to restrict the identities shown."

# Subset desired identities in a DimPlot.
p <- SCpubr::do_DimPlot(sample = sample[, sample$seurat_clusters %in% c("0", "5", "2", "4")])

p
```

However, we end up losing the **UMAP silhouette**. For this use case, `SCpubr::do_DimPlot()` introduces `idents.keep` parameter, for which you can provide a vector with the identities you want to keep. This will assign to the rest of the cells a value of `NA` and they will be colored according to `na.value` parameter:

```{r}
#| fig-width: 20
#| fig-height: 11

# Select identities with idents.keep.
p1 <- SCpubr::do_DimPlot(sample = sample,
                         idents.keep = c("0", "5", "2",  "4"))

# Also, non-selected cells's color can be modified.
p2 <- SCpubr::do_DimPlot(sample = sample,
                         idents.keep = c("0", "5", "2",  "4"),
                         na.value = "grey50")
p <- p1 | p2
p
```

## Group by another metadata variable

So far, all the DimPlots displayed have shown the identities as the ones that are currently set in the object. This can be consulted by using `Seurat::Idents(sample)`. However, naturally, we might want to display different metadata variables. This can be easily achieved by using `group.by` parameter:

```{r}
#| fig-width: 20
#| fig-height: 11

# Generate another metadata variable to group the cells by.
sample$annotation <- sample(c("A", "B", "C"), ncol(sample), replace = TRUE)

# Group by another metadata variable.
p1 <- SCpubr::do_DimPlot(sample, 
                         group.by = "seurat_clusters")

p2 <- SCpubr::do_DimPlot(sample, 
                         group.by = "annotation")

p <- p1 | p2
p
```

## Splitting by a category

Another useful parameter is `split.by`, which allows you to split your DimPlot into multiple panels, each one containing a different unique value of the metadata variable you have provided to the argument. One can understand this as using the `group.by` parameter and then splitting the resulting DimPlot into different panels. In this example, we are going to use the different clusters as an example. This is how it looks by default:


```{r}
#| fig-width: 25
#| fig-height: 10

# SCpubr's DimPlot using split.by
p <- SCpubr::do_DimPlot(sample, 
                        split.by = "seurat_clusters", 
                        ncol = 5, 
                        legend.position = "none",
                        font.size = 24)

p
```

If we are interested only in a subset of the possible values, we can use `idents.keep` alongside a vector containing the values to keep from the unique values in `split.by`:

```{r}
#| fig-width: 21
#| fig-height: 8

# Using split.by and restricting the number of output plots with idents.keep.
p <- SCpubr::do_DimPlot(sample, 
                        split.by = "seurat_clusters", 
                        ncol = 3, 
                        idents.keep = c("0", "1", "7"),
                        legend.position = "none",
                        font.size = 24)

p
```

## Group by a variable but split by another

Finally, but also importantly, users might want to **split** the UMAP using `split.by`, while also **grouping** (coloring) the values by another variable using `group.by`. Using these two parameters in combination yields the following:

```{r}
#| fig-width: 30
#| fig-height: 11

# Using split.by and group.by in combination.
sample$orig.ident <- sample(c("A", "B", "C"), 
                            ncol(sample), 
                            replace = TRUE, 
                            prob = c(0.05, 0.1, 0.85))

p <- SCpubr::do_DimPlot(sample, 
                        group.by = "seurat_clusters",
                        split.by = "orig.ident", 
                        font.size = 24)

p
```
