# Installation guide {-}

This package can be installed using the following commands:
```{r, eval = F}
# From CRAN:
# Future sumission to CRAN.

# From GitHub.
if(!requireNamespace("devtools", quietly = T)){
  install.packages("devtools") # If not installed.
}
devtools::install_github("enblacar/SCpubr")

```


Further packages are needed in order to run the different functions:

- [circlize](https://jokergoo.github.io/circlize_book/book/)
- [colorspace](https://cran.r-project.org/web/packages/colorspace/vignettes/colorspace.html)
- [ComplexHeatmap](https://bioconductor.org/packages/release/bioc/html/ComplexHeatmap.html)
- [dplyr](https://dplyr.tidyverse.org/)
- [enrichR](https://cran.r-project.org/web/packages/enrichR/vignettes/enrichR.html)
- [forcats](https://forcats.tidyverse.org/index.html)
- [ggbeeswarm](https://cran.r-project.org/web/packages/ggbeeswarm/index.html)
- [ggExtra](https://cran.r-project.org/web/packages/ggExtra/vignettes/ggExtra.html)
- [ggplot2](https://ggplot2.tidyverse.org/)
- [ggplotify](https://cran.r-project.org/web/packages/ggplotify/index.html/)
- [ggrastr](https://cran.r-project.org/web/packages/ggrastr/vignettes/Raster_geoms.html)
- [ggrepel](https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html)
- [ggridges](https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html)
- [ggsankey](https://github.com/davidsjoberg/ggsankey)
- [liana](https://github.com/saezlab/liana)
- [maggritr](https://cran.r-project.org/web/packages/magrittr/vignettes/magrittr.html)
- [monocle3](https://cole-trapnell-lab.github.io/monocle3/)
- [Nebulosa](http://bioconductor.org/packages/release/bioc/html/Nebulosa.html)
- [patchwork](https://patchwork.data-imaginist.com/)
- [pbapply](https://cran.rstudio.com/web/packages/pbapply/index.html)
- [plyr](https://cran.r-project.org/web/packages/plyr/index.html)
- [rlang](https://cran.r-project.org/web/packages/rlang/index.html)
- [scales](https://cran.r-project.org/web/packages/scales/index.html)
- [scattermore](https://cran.rstudio.com/web/packages/scattermore/index.html)
- [Seurat](https://satijalab.org/seurat/index.html)
- [stringr](https://cran.r-project.org/web/packages/stringr/index.html)
- [svglite](https://cran.r-project.org/web/packages/svglite/index.html)
- [tibble](https://tibble.tidyverse.org/)
- [tidyr](https://tidyr.tidyverse.org/)
- [viridis](https://cran.r-project.org/web/packages/viridis/index.html)


All packages can be installed using the following commands:

```{r, eval = F}
# Install CRAN packages.
cran_packages <- c("circlize",
                   "colorspace",
                   "dplyr",
                   "enrichR",
                   "forcats",
                   "ggbeeswarm",
                   "ggdist",
                   "ggExtra",
                   "ggplot2",
                   "ggplotify",
                   "ggrastr",
                   "ggrepel",
                   "ggridges",
                   "grDevices",
                   "grid",
                   "magrittr",
                   "patchwork",
                   "pbapply",
                   "plyr",
                   "rlang",
                   "scales",
                   "scattermore",
                   "Seurat",
                   "stats",
                   "stringr",
                   "svglite",
                   "tibble",
                   "tidyr",
                   "viridis")

install.packages(cran_packages)

# Install bioconductor packages.
bioconductor_packages <- c("ComplexHeatmap",
                           "infercnv",
                           "Nebulosa")
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install(bioconductor_packages)

# Install github packages.
github_packages <- c("ggsankey",
                     "liana",
                     "monocle3")

if (!requireNamespace("remotes", quietly = TRUE))
    install.packages("remotes")

remotes::install_github(github_packages)
```

One can check the dependencies by running `SCpubr::state_dependencies()`.

