# Dim plots

```{r echo = F}
sample <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/sc_dataset.rds")
```

Dimensional reduction Plots (**DimPlots**) are, probably, one of the most iconic visualizations from `Seurat`. It allows the user to visualize the cells in a dimensional reduction embedding such as `PCA` or `UMAP`. The cells can be, then, colored by any desired groups. In short, this visualization allows the user to plot **any kind of categorical data** onto the cells in the dimensional reduction embedding. 

## Basic usage

This is SCpubr's take on `Seurat::DimPlot()`. 

```{r, fig.cap = "SCpubr DimPlot vs Seurat DimPlot.", fig.width=14, fig.height=8}
# Seurat's DimPlot.
p1 <- Seurat::DimPlot(sample)

# SCpubr's DimPlot.
p2 <- SCpubr::do_DimPlot(sample = sample)

p <- p1 | p2
p
```

Even though axes are removed from UMAP reductions, the title for the axes is kept for any other reduction used. The same is applied for UMAP if the default dimension order is altered.

```{r, fig.cap = "SCpubr DimPlot with PCA embedding showing axes titles and UMAP changing default dimension component order.", fig.width=14, fig.height=8}
# Example using PCA reduction.
p1 <- SCpubr::do_DimPlot(sample = sample, 
                         reduction = "pca")

# Example using a non-canonical set of dimensions.
p2 <- SCpubr::do_DimPlot(sample = sample, 
                         dims = c(2, 1))

p <- p1 | p2
p
```

We can change the legend's number of columns or rows with `legend.ncol` and `legend.nrow`.

```{r, fig.cap = "SCpubr DimPlot with legend to the left.", fig.width=14, fig.height=8}
# Modify the number of columns in the legend.
p1 <- SCpubr::do_DimPlot(sample = sample, 
                         legend.ncol = 2)

# Modify the number of rows in the legend.
p2 <- SCpubr::do_DimPlot(sample = sample, 
                         legend.nrow = 3)

p <- p1 | p2 
p
```

## Using labels instead of a legend

In some cases, especially early on in the analysis where we do only have numbers as cluster names, we might want to remove the legend entirely, and instead plot labels on top of each cluster. This is achieved by using `label = TRUE`. This is how it looks in Seurat:

```{r, fig.cap = "Seurat DimPlot with labels on top of the clusters.", fig.width=7, fig.height=8}
# Put labels on top of the clusters.
p <- Seurat::DimPlot(sample, 
                     label = TRUE)
p
```    

However, we can play further with other parameters of the function such as `label.color` which will provide a different color for the text inside the labels:

```{r, fig.cap = "Seurat DimPlot with labels on top of the clusters and a different color.", fig.width=7, fig.height=8}
# Change the color of the text in the labels.
p <- SCpubr::do_DimPlot(sample = sample, 
                        label = TRUE, 
                        label.color = "black")
p
``` 

Finally, one can always get rid of the legend with `legend.position = "none`:

```{r, fig.cap = "Seurat DimPlot without legend.", fig.width=7, fig.height=8}
# Remove the legend from the plot.
p <- SCpubr::do_DimPlot(sample = sample, 
                        label = TRUE, 
                        legend.position = "none")
p
``` 

## Changing the order of plotting

By default, cells in `SCpubr::do_DimPlot()` are randomly plotted by using `shuffle = TRUE`. This is done as the default behavior of `Seurat::DimPlot()` is to plot the cells based on the factor levels of the identities. Sometimes, this way of plotting results in some clusters not being visible as another one is on top of it. However, this behavior might be intended, and even more, we would like a **specific** identity to stand out from the rest. This can be achieved by providing to `order` parameter either a vector with all the identities ordered however you want, or just some of the identities, which will then be plotted on top of the rest. However, it is still not clear whether using `order` alongside `shuffle = TRUE` have unexpected side effects. For this, please use it alongside `shuffle = FALSE`. For the following example, we are going to test:

- We are going to bring cluster 5 to the front.
- `order` with one value or all values.
- `order` with `shuffle = TRUE/FALSE`.
- We will increase the dot size to better see cells in cluster 7, as this data set has a lack of overlapping clusters..

```{r, fig.cap = "SCpubr, modifying order of plotted identities in a DimPlot", fig.width=14, fig.height=16}
# Regular SCpubr DimPlot.
p1 <- SCpubr::do_DimPlot(sample = sample,
                         reduction = "pca",
                         plot.title = "Normal DimPlot")

# Using order with one value and shuffle = TRUE.
p2 <- SCpubr::do_DimPlot(sample = sample,
                         shuffle = TRUE,
                         order = "5",
                         reduction = "pca",
                         plot.title = "shuffle = TRUE")

# Using order with one value and shuffle = FALSE.
p3 <- SCpubr::do_DimPlot(sample = sample,
                         shuffle = FALSE,
                         order = "5",
                         reduction = "pca",
                         plot.title = "shuffle = FALSE")

# Using order with all values.
p4 <- SCpubr::do_DimPlot(sample = sample,
                         shuffle = FALSE,
                         order = c("5", "8", "4",
                                   "9", "3", "1",
                                   "6", "0", "7", "2"),
                         reduction = "pca",
                         plot.title = "shuffle = FALSE all identities")

p <- (p1 | p2) / (p3 | p4)
p
```
We can see that cluster 5 always plots on top of cluster 0 when `order` is set. While still not clear how `order` and `shuffle` interact, it is apparent that using `order` cancels the original behavior of `shuffle`. Therefore, if both `order` and `shuffle` are set, `SCpubr::do_DimPlot()` will throw a warning.


## Highlighting cells

One of the nice features of `Seurat::DimPlot()` is the possibility of highlighting a certain group of cells in the plot. This is achieved by using the `cells.highligh` parameter. This is how the default plot looks like and `SCpubr::do_DimPlot()`'s take on it:

```{r, fig.cap = "Highlighting cells, comparison between Seurat and SCpubr.", fig.width=14, fig.height=8}
cells.use <- sample(x = colnames(sample), 
                    size = 1500)

# Compare Seurat and SCpubr way of highlighting cells.
p1 <- Seurat::DimPlot(sample, 
                      cells.highlight = cells.use)

p2 <- SCpubr::do_DimPlot(sample = sample,
                         cells.highlight = cells.use)

p <- p1 | p2
p
```

One can also change the color of the highligted cells by providing single color to `colors.use` and the color of the not selected cells with `na.value`:

```{r, fig.cap = "SCpubr DimPlot highlighting cells and changing default colors.", fig.width=7, fig.height=8}
# Change color of highlighted and non-highlighted cells.
p <- SCpubr::do_DimPlot(sample = sample, 
                        cells.highlight = cells.use,
                        colors.use = "black",
                        na.value = "grey90")
p
```

By default, the size of all cells in `SCpubr::do_DimPlot()` is the same. However, the size of the highlighted dots can be modified with the parameter `sizes.highlight`.

```{r, fig.cap = "SCpubr DimPlot highlighting cells bigger dot size.", fig.width=7, fig.height=8}
# Increase the size of the highlighted cells.
p <- SCpubr::do_DimPlot(sample = sample, 
                        cells.highlight = cells.use, 
                        sizes.highlight = 1)
p
```
We can also highlight **whole identities** with `idents.highlight` parameter. For this, just provide the desired identities to be selected. It can also work in combination with `cells.highlight`.

```{r, fig.cap = "SCpubr DimPlot highlighting cells testing combinations of cells.highlight and idents.highlight.", fig.width=21, fig.height=8}
# Using cells.highlight.
p1 <- SCpubr::do_DimPlot(sample = sample, 
                         cells.highlight = cells.use)

# Using idents.highlight.
p2 <- SCpubr::do_DimPlot(sample = sample, 
                         idents.highlight = c("6"))

# Using both.
p3 <- SCpubr::do_DimPlot(sample = sample, 
                         cells.highlight = cells.use, 
                         idents.highlight = c("6"))

p <- p1 | p2 | p3
p
```

## Restrict the identitites shown and grey out the rest
Sometimes, we are interested in showing only some of the identities or groups in our sample, but instead of the previous approach, we still want to keep the original colors and legend. For this, we might be inclined to just subset this sample, as follows:
```{r, fig.cap = "Seurat DimPlot selecting only some identities by subsetting the sample.", fig.width=7, fig.height=8}
# Subset desired identities in a DimPlot.
p <- SCpubr::do_DimPlot(sample = sample[, sample$seurat_clusters %in% c("0", "5")])

p
``` 

However, we end up losing the **UMAP silhouette**. For this, `SCpubr::do_DimPlot()` introduces `idents.keep` parameter, for which you can provide a vector with the identities you want to keep. This will assign to the rest of the cells a value of `NA` and they will be colored according to `na.value` parameter:

```{r, fig.cap = "Seurat DimPlot selecting only some identities by using idents.keep.", fig.width=14, fig.height=8}
# Select identities with idents.keep.
p1 <- SCpubr::do_DimPlot(sample = sample,
                         idents.keep = c("0", "5"))

# Also, non-selected cells's color can be modified.
p2 <- SCpubr::do_DimPlot(sample = sample,
                         idents.keep = c("0", "5"),
                         na.value = "grey50")
p <- p1 | p2
p
``` 

## Splitting by a category

Another useful parameter of `Seurat::DimPlot` is `split.by`, which allows you to split your DimPlot into multiple panels, each one containing a different unique value of the metadata variable you have provided to the argument. One can understand this as using the `group.by` parameter and then splitting the resulting DimPlot into different panels. In this example, we are going to use the different clusters as an example. This is how it looks by default:

```{r, fig.cap = "Seurat DimPlot using split.by.", fig.width=10, fig.height=5}
# Seurat's DimPlot using split.by
p <- Seurat::DimPlot(sample, 
                     split.by = "seurat_clusters", 
                     ncol = 5)

p
```
As can be observed, this plot accomplish the task of separating the cells into each panel, but the approach followed actually makes **interpretation difficult**. Clusters such as Cluster 9, with fewer cells, tell pretty much nothing. Besides, losing the **UMAP silhouette** is a major downside of this approach. This is where `SCpubr` focus on. It generates as many plots as unique values in the metadata to split the plot by, but uses `cells.highlight` instead, which keeps the rest of cells greyed out according to `na.value` color. This is how it looks:

```{r, fig.cap = "SCpubr DimPlot using split.by.", fig.width=25, fig.height=10}
# SCpubr's DimPlot using split.by
p <- SCpubr::do_DimPlot(sample, 
                        split.by = "seurat_clusters", 
                        ncol = 5, 
                        legend.position = "none",
                        font.size = 24)

p
```

This way, we can see that clusters such as Cluster 7 are way more disperse than the rest, accounting not only for standalone groups of cells but also blending into other bigger clusters. 

If we are interested only in a subset of the possible values, we can use `idents.keep` alongside a vector containing the values to keep from the unique values in `split.by`:

```{r, fig.cap = "SCpubr DimPlot using split.by and idents.keep.", fig.width=21, fig.height=8}
# Using split.by and restricting the number of output plots with idents.keep.
p <- SCpubr::do_DimPlot(sample, 
                        split.by = "seurat_clusters", 
                        ncol = 3, 
                        idents.keep = c("0", "1", "7"),
                        legend.position = "none",
                        font.size = 24)

p
```
## Group by a variable but split by another
Finally, but also importantly, users might want to **split** the UMAP using `split.by`, while also **grouping** (coloring) the values by another variable using `group.by`. Using these two parameters in combination yields the following:

```{r, fig.cap = "SCpubr DimPlot using split.by and group.by", fig.width=21, fig.height=8}
# Using split.by and group.by in combination.
sample$orig.ident <- sample(c("A", "B", "C"), ncol(sample), replace = TRUE, prob = c(0.05, 0.1, 0.85))

p <- SCpubr::do_DimPlot(sample, 
                        group.by = "seurat_clusters",
                        split.by = "orig.ident", 
                        font.size = 24)

p
```


Actually, the user might want to change the color of the highlighted cells in this split DimPlot. This is achieved by using `colors.use` parameter and providing either a vector of **valid color representations** of equal length to unique values in `split.by` or just a single color to use in all panels.


```{r, fig.cap = "SCpubr DimPlot using split.by with a changed color", fig.width=25, fig.height=20}
# Create a color scale for the unique values in seurat clusters.
colors <- c("0" = "#001219",
            "1" = "#005f73",
            "2" = "#0a9396",
            "3" = "#94d2bd",
            "4" = "#e9d8a6",
            "5" = "#ee9b00",
            "6" = "#ca6702",
            "7" = "#bb3e03",
            "8" = "#ae2012",
            "9" = "#9b2226")

# Use a custom set of colors, one for each identity.
p1 <- SCpubr::do_DimPlot(sample, 
                         split.by = "seurat_clusters", 
                         ncol = 5, 
                         legend.position = "none", 
                         colors.use = colors,
                         font.size = 24)

# Use the same color for all identities.
p2 <- SCpubr::do_DimPlot(sample, 
                         split.by = "seurat_clusters", 
                         ncol = 5, 
                         legend.position = "none",
                         colors.use = "#008080",
                         font.size = 24)
                         
p <- p1 / p2
p
```



