# Nebulosa plots

This plot comes straight from the [Nebulosa package](http://bioconductor.org/packages/release/bioc/html/Nebulosa.html). For more information, here is their [vignette](http://bioconductor.org/packages/release/bioc/vignettes/Nebulosa/inst/doc/nebulosa_seurat.html). In short, it computes a "density" of whether the surrounding cells (in the UMAP embedding) also express the variable that you provide to the function, same variables that one could feed to `Seurat::FeaturePlot()`. This is how their default plot looks like:



```{r echo = F}
sample <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/sc_dataset.rds")
```

## Basic usage
```{r, fig.cap= "Nebulosa plot", fig.width=8, fig.height=7}
p <- Nebulosa::plot_density(sample,
                            features = "CD14")
p
```

This way, not only we do know which clusters are enriched in CD14, we also know the regions of it with the highest density of cells expressing it. Minor modifications, on the line of the rest of plots, are applied in `SCpubr`:
```{r, fig.cap = "SCpubr Nebulosa plot.", fig.width=7, fig.height=8}
p <- SCpubr::do_NebulosaPlot(sample = sample, 
                             features = "CD14")
p
```

Then, this type visualization becomes a natural partner to `Seurat::FeaturePlot()' as not only we are able to visualize the expression of a variable, but also query the density of the surrounding cells. Here is an example:

```{r, fig.cap = "SCpubr FeaturePlot and NebulosaPlot.", fig.width=14, fig.height=8}
p1 <- SCpubr::do_FeaturePlot(sample = sample, 
                             features = "CD14") 
p2 <- SCpubr::do_NebulosaPlot(sample = sample, 
                              features = "CD14")
p1 | p2
```

[Nebulosa](https://github.com/powellgenomicslab/Nebulosa/blob/master/R/plotting.R) package also offers the option to query multiple features at the same time:

```{r, fig.cap = "SCpubr Nebulosa plot with multiple features.", fig.width=14, fig.height=8}
p <- SCpubr::do_NebulosaPlot(sample, 
                             features = c("CD14", "CD8A"))
p
```

## Compute joint densities
But, more intriguingly, it can also compute the **joint** density of the features. This is achieved by parsing `joint = TRUE`.

```{r, fig.cap = "SCpubr Nebulosa plot with joint density", fig.width=21, fig.height=8}
p <- SCpubr::do_NebulosaPlot(sample = sample, 
                             features = c("CD14", "CD8A"), 
                             joint = TRUE)
p 
```

If we were interested in retrieving only the joint density plot, we can accomplish it with `return_only_joint` parameter. Please note that, since this will return only one plot, if wanted to modify the plot title, use `plot.title` instead:

```{r, fig.cap = "SCpubr Nebulosa plot with joint density, selecting only the last plot", fig.width=7, fig.height=8}
features.use <- c("CD14", "CD8A")

p <- SCpubr::do_NebulosaPlot(sample = sample, 
                             features = c("CD14", "CD8A"), 
                             joint = TRUE, 
                             return_only_joint = TRUE,
                             plot.title = "Joint density CD14+-CD8A+")

p
```
## Modify titles individually
If we query multiple features or use `joint = TRUE`, then using `plot.title` will add a general title to the different panels generated. To then modify the plot title of each individual panel, a vector **of the same length as the number of features** needs to be provided to `individual.titles` parameter. In case a title does not need to be changed, use `NA` instead. Please take note that `joint = TRUE` will generate an extra panel that needs another plot title. 

```{r, fig.cap = "SCpubr Nebulosa plot with joint density, modifying each individual plot's name", fig.width=21, fig.height=8}
p <- SCpubr::do_NebulosaPlot(sample = sample, 
                             features = c("CD14", "CD8A"), 
                             joint = TRUE, 
                             individual.titles = c("Plot A",
                                                   NA,
                                                   "Combined density"),
                             plot.title = "Density analysis")
p
```

