# Box plots

```{r echo = F}
sample <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/sc_dataset.rds")
```

Deep intro goes here.

## Basic usage

To generate a basic box plot you need to provide the seurat object and a feature to plot:

```{r, fig.cap = "Basic box plot.", fig.width=10, fig.height=7}
# Basic box plot.
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA")
p
```

## Use custom grouping
By default, the groups are the identities in the seurat object, but this can be controlled using `group.by parameter`.

```{r, fig.cap = "Custom groups.", fig.width=10, fig.height=7}
# Generate a custom group.
sample$custom_group = ifelse(colnames(sample) %in% sample(colnames(sample), 4000), "A", "B")

# Use custom grouping.
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA",
                        group.by = "custom_group")
p
```

We can flip the plot if needed with `flip = TRUE`:

```{r, fig.cap = "Flip the box plot.", fig.width=7, fig.height=10}
# Flip the box plot.
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA",
                        flip = TRUE)
p
```

## Modify aesthetic style
We can also apply another aesthetic style, which just colors the lines of the boxplots and leave the rest white with `use_silhouette = TRUE`:

```{r, fig.cap = "Apply silhouette style.", fig.width=10, fig.height=7}
# Use silhouette style.
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA",
                        use_silhouette = TRUE)
p
```

## Reorder by mean values
We can reorder the values by their average mean, from highest to lowest with `order = TRUE`:

```{r, fig.cap = "Order by mean.", fig.width=10, fig.height=7}
# Order by mean values.
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA",
                        order = TRUE)
p
```

## Split by a second variable
We can further apply a second grouping by using `split.by`.

```{r, fig.cap = "Apply a second grouping.", fig.width=10, fig.height=7}
# Apply second grouping.
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA",
                        split.by = "custom_group")
p
```

## Apply statistical tests to compare groups
Finally we can apply statistical tests to the groups, if we do not use `split.by`. We can do that by setting `use_test = TRUE` and providing the comparison as a list of pairs of items. The test, by default is wilcoxon test, but can be modified using `test`. *The correct use of such tests and their interpretation relies on the user*.

```{r, fig.cap = "Apply statistical tests.", fig.width=10, fig.height=7}
# Apply statistical tests.
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA",
                        use_test = TRUE,
                        comparisons = list(c("0", "1"),
                                           c("3", "4"),
                                           c("5", "9")))
p
```

One can display this on symbols, or with numbers with `map_signif_level`:

```{r, fig.cap = "Apply statistical tests and show the p-value", fig.width=10, fig.height=7}
# Apply statistical tests and show the p-value.
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA",
                        use_test = TRUE,
                        comparisons = list(c("0", "1"),
                                           c("3", "4"),
                                           c("5", "9")),
                        map_signif_level = FALSE)
p
```
