# Volcano plots

Perhaps one of the most known type of plots for bulk transcriptomics. By computing DE genes across two conditions, the results can be plotted as a volcano plot. This plot features the genes as dots, and places them in a scatter plot where the X axis contains the degree in which a gene is differentially expressed (average log2(FC)), while the Y axis shows the how significant the gene is (-log10(p-value adjusted)). 

## Basic usage
To generate such a plot, one can use `SCpubr::do_VolcanoPlot()`, which needs as input the Seurat object and the result of running `Seurat::FindMarkers()` choosing two groups.

```{r echo = F}
sample <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/sc_dataset.rds")
de_genes <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/pairwise_de_genes.rds")
```

```{r, fig.cap="SCpubr::do_VolcanoPlot with default parameters.", fig.width=10, fig.height=10}
# Generate a volcano plot.
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes)
p
```

As you can see, there are four major groups of genes:
- Genes that surpass our p-value and logFC cutoffs (blue).
- Genes that surpass the p-value cutoff but not the logFC cutoff (orange).
- Genes that surpass the logFC cutoff but not the p-value cutoff (purple, not shown).
- Genes that do not surpass any cutoff (green).

## Modify the cutoffss

The cutofss can be set up by the user using `pval_cutoff` (without -log10 transforming) and `FC_cutoff` (avg log2(FC)).

```{r, fig.cap="SCpubr::do_VolcanoPlot modifying cutoffs.", fig.width=10, fig.height=10}
# Modify cutoffs.
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes,
                            pval_cutoff = 1e-15,
                            FC_cutoff = 0.2)
p
```

## Modify the gene tags
By default, the top 5 genes on each side, ordered by `-log10(p-value adjusted)` and `average log2(FC)` are reported. However, one can increase the number by using

```{r, fig.cap="SCpubr::do_VolcanoPlot modifying the number of gene tags.", fig.width=10, fig.height=10}
# Modify number of gene tags.
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes,
                            n_genes = 15)
p
```








