# Sankey plots

::: callout-important
## Please note:

`SCpubr::do_SankeyPlot()` is discontinued and will be removed from SCpubr in the following version. Please consider using `SCpubr::do_AlluvialPlot()` instead.
:::

::: callout-warning
## Please note:

Due to CRAN policies, this function could not make it to the **official release** and is only available in the **development release** of **SCpubr**. Find how to install it [**here**](https://github.com/enblacar/SCpubr/releases/tag/v1.1.2-dev-stable).
:::

```{r}
#| results: "asis"
#| echo: false
source("_common.R")
```

```{r}
#| echo: false
sample <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/sc_dataset.rds")
library(dplyr)
```

Sankey and alluvial plots are a set of visualizations that help depicting how the cells "flow" from a given group to another. This can be particular interesting, let's say, when we have a merged dataset that has samples from different subtypes and we want to see how the cells from each subtype distribute along the different clusters in our sample. Or, similarly, how the cells from each cluster distribute across the different individual datasets.

## Basic usage

**Due to an internal bug in ggsankey, the package dplyr has to be loaded with `library(dplyr)` before running this function.**

This is how we can achieve basic Sankey or alluvial plots (using \link[ggsankey]{https://github.com/davidsjoberg/ggsankey} package):

```{r}
#| fig-width: 7
#| fig-height: 14
#| fig-cap: "Basic Sankey and Alluvial plots."

# Generate a more fine-grained clustering.
sample <- Seurat::FindSubCluster(sample, cluster = c("0", "5"), graph.name = "SCT_snn")
sample$sub.cluster <- paste0("sub_", sample$sub.cluster)

# Compute basic sankey plot.
p1 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            last_group = "seurat_clusters",
                            type = "sankey")

# Compute basic alluvial plot.
p2 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            last_group = "seurat_clusters",
                            type = "alluvial")

p <- p1 / p2
p
```

The key aspects of this figure is that one has to provide the starting and end groups to `first_group` and `last_group` parameters, and the type of plot to `type`, making it either alluvial or sankey. The major difference between both is simply that the alluvial plot has not space between the groups. This allows for a more compact visualization, although it can get quite messy depending on the colors. We can keep adding as many groups as we want in between by using `middle_groups` parameter.

```{r}
#| fig-width: 7
#| fig-height: 14
#| fig-cap: "Add more groups in the middle."

sample$assignment <- ifelse(sample$seurat_clusters %in% c("0", "2", "4"), "A", "B")

# Add more groups.
p1 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey")

p2 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "alluvial")

p <- p1 / p2
p
```

## Modify node aesthetics

By default, the color of the nodes and the fill is white, so that they can not be seen. One Can modify this by using `node.fill` and `node.color`.

```{r}
#| fig-width: 7
#| fig-height: 14
#| fig-cap: "Modify the colors for the nodes."

# Control the color and fill of the nodes.
p1 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey",
                            node.fill = "grey95",
                            node.color = "black")

p2 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "alluvial",
                            node.fill = "grey95",
                            node.color = "black")

p <- p1 / p2
p
```

As can be seen, the text labels, then, overlap the limits of the nodes. We can fix that using the `width` parameter:

```{r}
#| fig-width: 7
#| fig-height: 14
#| fig-cap: "Modify the width of the nodes."

# Control the width of the nodes.
p1 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey",
                            node.fill = "grey95",
                            node.color = "black")

p2 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey",
                            node.fill = "grey95",
                            node.color = "black",
                            width = 0.5)

p <- p1 / p2
p
```

## Control alignment of the labels

We can also control the alignment of the labels (goes different by the different groups by default) using the `hjust` parameter. This will modify it for all the nodes.

```{r}
#| fig-width: 7
#| fig-height: 14
#| fig-cap: "Control the alignment of the labels."

# Control the alignment of the labels.
p1 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey",
                            node.fill = "grey95",
                            node.color = "black",
                            width = 0.5)

p2 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey",
                            node.fill = "grey95",
                            node.color = "black",
                            width = 0.5,
                            hjust = 0.5)

p <- p1 / p2
p
```

We can also turn the text into labels, which will then be colored based on the end nodes. We can toggle it by using `use_labels = TRUE`.

```{r}
#| fig-width: 7
#| fig-height: 14
#| fig-cap: "Use text or labels."

# Use text or labels for the nodes.
p1 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey")

p2 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey",
                            use_labels = TRUE,
                            text_color = "white")

p <- p1 / p2
p
```

## Increase the gap between the nodes

One can also increase the gaps between the nodes. This is done by providing a number to `space`. This number would act as "invisible" cells in between the nodes. So one needs to think in terms of how many cells do we have in our dataset. The bigger the dataset, the higher the number we need to notice a difference. It is set by default to 5% of the total number of cells in the dataset.

```{r}
#| fig-width: 7
#| fig-height: 14
#| fig-cap: "Increase the gaps between the nodes."

# Modify the space between nodes.
p1 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey",
                            space =  1000)

p2 <- SCpubr::do_SankeyPlot(sample = sample,
                            first_group = "sub.cluster",
                            middle_groups = c("seurat_clusters", "assignment"),
                            last_group = "orig.ident",
                            type = "sankey",
                            space = 5000)

p <- p1 / p2
p
```

## Modify default colors

Finally, we can modify the default colors by providing our own to `colors.first`, `colors.middle` and `colors.last`. It has to be a named vector with

```{r}
#| fig-width: 7
#| fig-height: 7
#| fig-cap: "Change default colors."

# Modify default colors.
colors.first <- SCpubr::do_ColorPalette(colors.use = "steelblue",
                                        n = length(unique(sample$sub.cluster)))
names(colors.first) <- unique(sample$sub.cluster)

p <- SCpubr::do_SankeyPlot(sample = sample,
                           first_group = "sub.cluster",
                           middle_groups = c("seurat_clusters", "assignment"),
                           last_group = "orig.ident",
                           type = "sankey",
                           colors.first = )
p
```
