# Enrichment score heatmaps

A very common approach to make sense of your cells is to query several list of marker genes, retrieved from literature, and compute how enriched each cell is in each given list of genes. This is achieved by using `Seurat::AddModuleScore`. The scores can be then visualized as a Feature plot, but one can also aggregate the enrichment scores by any variable of interest, for instance the different clusters in the sample.

```{r}
#| results: "asis"
#| echo: false
source("_common.R")
suppressMessages(library("Seurat"))
```

```{r}
#| echo: false
sample <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/sc_dataset.rds")
sample$annotation <- sample(c("A", "B"), ncol(sample), replace = TRUE)
```

This kind of heatmaps can be easily computed using `SCpubr::do_EnrichmentHeatmap()`:

## Single grouping variable

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Basic output."

# Define list of genes.
genes <- list("Naive CD4+ T" = c("IL7R", "CCR7"),
              "CD14+ Mono" = c("CD14", "LYZ"),
              "Memory CD4+" = c("S100A4"),
              "B" = c("MS4A1"),
              "CD8+ T" = c("CD8A"),
              "FCGR3A+ Mono" = c("FCGR3A", "MS4A7"),
              "NK" = c("GNLY", "NKG7"),
              "DC" = c("FCER1A", "CST3"),
              "Platelet" = c("PPBP"))

# Default parameters.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  viridis_direction = -1)
p
```

::: callout-note
## Please note:

In these cases, inverting the color scale with `viridis_direction = -1` is the best choice as makes it easir to identify dark colors in the heatmap and they stand out more than it would do when it comes to using light colors on dark background.
:::

By default, `SCpubr::do_EnrichmentHeatmap` aggregates the values by the current identity. However, other metadata variables can be used to aggregate for. For this, provide the name to `group.by` parameter.

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Group by a custom variable."

# Custom aggregated values.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  group.by = "orig.ident",
                                  viridis_direction = -1)
p
```

## Several groupoing variables

However, more than one variable can be passsed at the same time to `group.by`:

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Group by several variables."

# Group by several variables.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  group.by = c("seurat_clusters", "orig.ident", "annotation"),
                                  viridis_direction = -1)
p
```

One can also customize the group titles by providing as many characters as variables in `group.by` to `row_title`:

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Group by a custom variable."

# Custom aggregated values.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  group.by = c("seurat_clusters", "orig.ident", "annotation"),
                                  row_title = c("A", "B", "C"),
                                  viridis_direction = -1)
p
```

## Transpose the heatmaps

The heatmaps can be transposed using `flip = TRUE`.

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Transpose the matrix."

# Transposing the matrix.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  flip = TRUE,
                                  viridis_direction = -1)
p
```

## Modify the rotation of row and column titles

Both rows and column names can be rotated using `column_names_rot` and `row_names_rot` parameters, providing the desired angle.

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Rotate the labels."

# Rotating the labels.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  flip = TRUE,
                                  column_names_rot = 0,
                                  viridis_direction = -1)
p
```

## Changing the cell size in the heatmap.

By design, the aspect ratio of the tiles in the heatmap is fixed so that cells are squares, and not rectangles. However, the user has the possibility to increase/decrease the cell size of each tile by modifying `cell_size` parameter. This is set to 5 by default.

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Change the size of the cells."

# Modifying the tile size.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  flip = TRUE,
                                  cluster_cols = FALSE,
                                  cluster_rows = TRUE,
                                  cell_size = 7,
                                  viridis_direction = -1)
p
```

## Symmetrical scales

If one wants a symmetrical scale,

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Use a symmetrical color scale."

# Symmetrical scale viriis.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  flip = TRUE,
                                  cluster_cols = FALSE,
                                  cluster_rows = TRUE,
                                  enforce_symmetry = TRUE)
p
```

For such cases, it is best to set `use_viridis = FALSE`.

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Use a symmetrical color scale that is not viridis."

# Modifying the symmetrical scale non viridis.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  flip = TRUE,
                                  cluster_cols = FALSE,
                                  cluster_rows = TRUE,
                                  enforce_symmetry = TRUE,
                                  use_viridis = FALSE)
p
```

## Compute other type of scorings

Similar to `Seurat::AddModuleScore()`, other enrichment scoring methods have been developed. One of such is [**UCell**](https://github.com/carmonalab/UCell). To use it as scoring method, simply pass `flavor = "UCell"`:

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Use UCell Scoring."

# Use UCell scoring.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  flip = TRUE,
                                  cluster_cols = FALSE,
                                  cluster_rows = TRUE,
                                  flavor = "UCell",
                                  viridis_direction = -1)
p
```

Alternatively, [**AUCell**](https://bioconductor.org/packages/release/bioc/html/AUCell.html) scoring is also available using `flavor = "AUCell"`.

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Use AUCell Scoring."

# Use AUCell scoring.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  flip = TRUE,
                                  cluster_cols = FALSE,
                                  cluster_rows = TRUE,
                                  flavor = "AUCell",
                                  viridis_direction = -1)
p
```

## Subset the color scale.

It might be the case that a single gene set, for a single cluster, drives the color scale entirely. We can treat this as an outlier, so that we can subset the color scale to better represent the values in between. We can achieve that by using `min.cutoff` and/or `max.cutoff`.

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Subset the color scale."

# Subset the color scale.
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = genes,
                                  flip = TRUE,
                                  cluster_cols = FALSE,
                                  cluster_rows = TRUE,
                                  flavor = "UCell",
                                  min.cutoff = 0.2,
                                  max.cutoff = 0.7,
                                  viridis_direction = -1)
p
```
