# Pathway Activity inference analysis

**This section is highly similar to TF Activity inference analysis as it makes use of the same source package and visualizations.**

Another very important analysis that can be carried out, once we have our cells defined into groups (i.e: cell clusters), is Pathway Activity inference analysis. Out of the different tools out there to perform this task, the one we will use is called [decoupleR](https://github.com/saezlab/decoupleR) ([P.Badia-i-Mompel *et al*. 2022](https://academic.oup.com/bioinformaticsadvances/article/2/1/vbac016/6544613?login=true)). This tool allows the inference of biological activities out of Omics data. For this, it requires a dataset to act as prior knowledge. For Pathway Activity inference, [progeny](https://github.com/saezlab/progeny/) is used. This allows for the computation of activity scores on a cell basis depicting how much (or little) each cell is enriched in each of the cancer Pathways stored in the database.

In order to visualize the enrichment of our cells in the pathways, the results need to be computed:

```{r echo = F}
sample <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/sc_dataset.rds")
activities <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/progeny_scores_decoupleR.rds")
```

```{r, eval = F}
# Define your sample and assay.
sample <- your_seurat_object
assay <- "your_normalized_data_assay"

# Retrieve prior knowledge network.
network <- decoupleR::get_progeny(organism = "human")

# Run weighted means algorithm.
activities <- decoupleR::run_wmean(mat = as.matrix(sample@assays[[assay]]@data),
                                   network = network,
                                   .source = "source",
                                   .targe = "target",
                                   .mor = "weight",
                                   times = 100,
                                   minsize = 5)
```

With this, we can proceed to plot our results with `SCpubr::do_PathwayActivityPlot()`.

## Heatmap of averaged scores

The most informative and, perhaps, straightforward approach is to visualize the resulting scores averaged by the groups we have defined, as a heatmap. This is the default output of `SCpubr::do_PathwayActivityPlot()`.

```{r, fig.cap="SCpubr do_PathwayActivityPlot heatmap output.", fig.width=8, fig.height=7}
# General heatmap.
out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities)
p <- out$heatmaps$average_scores
p
```

## Feature plots of the scores

Perhaps we are interested into visualizing the scores as a Feature plot. This way we can observe trends of enrichment in the activities. This can be achieved by providing `plot_FeaturePlots = TRUE`.

```{r, fig.cap="SCpubr do_PathwayActivityPlot feature plot output.", fig.width=14, fig.height=8}
# Retrieve feature plots.
out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities,
                                      plot_FeaturePlots = TRUE)
p1 <- SCpubr::do_DimPlot(sample)
p2 <- out$feature_plots$EGFR
p <- p1 | p2
p

```

## Geyser plots of the scores

In the same fashion as with the feature plots, we can also visualize the scores as Geyser plots, to get a sense of the distribution of the scores alongside the groups. This can be achieved by providing `plot_GeyserPlots = TRUE`.

```{r, fig.cap="SCpubr do_PathwayActivityPlot geyser plot output.", fig.width=14, fig.height=8}
# Retrieve Geyser plots.
out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities,
                                      plot_GeyserPlots = TRUE)
p1 <- SCpubr::do_DimPlot(sample)
p2 <- out$geyser_plots$EGFR
p <- p1 | p2
p
```

## Use non-symmetrical color scales

If one wants to just plot a continuous color scale for the Feature plots and the Geyser plots, this can be achieved by using `enforce_symmetry = FALSE`.

```{r, fig.cap="SCpubr do_PathwayActivityPlot non-symmetrical color scale.", fig.width=14, fig.height=16}
# Use non-symmetrical color scale.
out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities,
                                      plot_GeyserPlots = TRUE,
                                      plot_FeaturePlots = TRUE,
                                      enforce_symmetry = FALSE)
p1 <- out$feature_plots$EGFR
p2 <- out$geyser_plots$EGFR

out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities,
                                      plot_GeyserPlots = TRUE,
                                      plot_FeaturePlots = TRUE,
                                      enforce_symmetry = TRUE)
p3 <- out$feature_plots$EGFR
p4 <- out$geyser_plots$EGFR

p <- (p1 | p2) / (p3 | p4)
p
```

## Order geysers by the mean

We can also decide not to order the Geyser plots by the mean of the values. We can do that by providing `geyser_order_by_mean = FALSE`.

```{r, fig.cap="SCpubr do_PathwayActivityPlot not ordering by mean values.", fig.width=14, fig.height=8}
# Not order Geyser plot by mean values.
out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities,
                                      plot_GeyserPlots = TRUE,
                                      enforce_symmetry = TRUE,
                                      geyser_order_by_mean = FALSE)
p1 <- out$geyser_plots$EGFR

out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities,
                                      plot_GeyserPlots = TRUE,
                                      enforce_symmetry = TRUE,
                                      geyser_order_by_mean = TRUE)
p2 <- out$geyser_plots$EGFR

p <- p1 | p2
p
```

## Plot another variable in the geyser plots

Furthermore, we can also recolor the Geyser plots based on another variable of our choice. We can achieve this by using `geyser_color.by` and `geyser_scale_type`, providing the name of the variable to plot and the type it is (continuous, categorical).

```{r, fig.cap="SCpubr do_PathwayActivityPlot using color.by.", fig.width=14, fig.height=8}
# Plot a third variable in Geyser plots.
out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities,
                                      plot_GeyserPlots = TRUE,
                                      geyser_color.by = "seurat_clusters",
                                      geyser_scale_type = "categorical")
p1 <- out$geyser_plots$EGFR

out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities,
                                      plot_GeyserPlots = TRUE,
                                      geyser_color.by = "nCount_RNA",
                                      geyser_scale_type = "continuous")
p2 <- out$geyser_plots$EGFR

p <- p1 | p2
p
```

## Split the heatmap into different groups

We can also further divide the heatmap into groups. This can be achieved using `split.by` parameter.

```{r, fig.cap="SCpubr do_PathwayActivityPlot splitting the heatmap.", fig.width=8, fig.height=7}
# Split the heatmap by another variable.
sample$split.me <- ifelse(sample$seurat_clusters %in% c("0", "3", "7"), "Group A","Group B")

out <- SCpubr::do_PathwayActivityPlot(sample = sample,
                                      activities = activities,
                                      split.by = "split.me")
p <- out$heatmaps$average_scores
p
```
