# TF Activity inference analysis

**This section is highly similar to Pathway Activity inference analysis as it makes use of the same source package and visualizations.**

Transcriptional Factor (TF) Activity inference analysis can also be carried out in a SC dataset. Same as with the previous chapter, we will use [decoupleR](https://github.com/saezlab/decoupleR) ([P.Badia-i-Mompel *et al*. 2022](https://academic.oup.com/bioinformaticsadvances/article/2/1/vbac016/6544613?login=true)). For this, [dorothea](https://github.com/saezlab/dorothea/) is used. This allows for the computation of activity scores on a cell basis depicting how much (or little) each cell is enriched in a TF and its dowstream targets (regulon).

In order to visualize the enrichment of our cells in the pathways, the results need to be computed:

```{r}
#| results: "asis"
#| echo: false
source("_common.R")
suppressMessages(library("Seurat"))
```

```{r}
#| echo: false
sample <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/sc_dataset.rds")
activities <- readRDS("/b06x-isilon/b06x-g/G703/eblanco/projects/test_SC_datasets/dorothea_scores_decoupleR.rds")
```

```{r, eval = F}
#| eval: false

# Define your sample and assay.
sample <- your_seurat_object
assay <- "your_normalized_data_assay"

# Retrieve prior knowledge network.
network <- decoupleR::get_dorothea(organism = "human",
                                   levels = c("A", "B", "C"))

# Run weighted means algorithm.
activities <- decoupleR::run_wmean(mat = as.matrix(sample@assays[[assay]]@data),
                                   network = network,
                                   .source = "source",
                                   .targe = "target",
                                   .mor = "mor",
                                   times = 100,
                                   minsize = 5)
```

With this, we can proceed to plot our results with `SCpubr::do_TFActivityPlot()`.

## Heatmap of averaged scores

The most informative and, perhaps, straightforward approach is to visualize the resulting scores averaged by the groups we have defined, as a heatmap. This is the default output of `SCpubr::do_TFActivityPlot()`.

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Retrieve the heatmap output."

# General heatmap.
out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities)
p <- out$heatmaps$average_scores
p
```

## Set the scale limits.

We can set the limits of the color scale by using `min.cutoff` and/or `max.cutoff`:

```{r}
#| fig-width: 12
#| fig-height: 7
#| fig-cap: "Set scale limits"

# Set the scale limits.
out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 min.cutoff = -2,
                                 max.cutoff = 2)
p <- out$heatmaps$average_scores
p
```

However, be mindful that if `enforce_symmetry = TRUE`, this will override these settings, to achieve a symmetrical color scale:

```{r}
#| fig-width: 12
#| fig-height: 7
#| fig-cap: "Effect of enforce_symmetry in min.cutoff and max.cutoff"

# Effect of enforce symmetry on the scale limits.
out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 min.cutoff = -2,
                                 max.cutoff = 1,
                                 enforce_symmetry = TRUE)
p <- out$heatmaps$average_scores
p
```

This effect will also be seen if only `min.cutoff` or `max.cutoff` is set and the other end of the color scale has a higher absolute value.

::: callout-note
## Take-home message

Be mindful when using `min.cutoff` and `max.cutoff`, as this behavior might be overriden by `enforce_symmetry = TRUE`. Consider setting it to `FALSE`.
:::

## Select the number of top TFs to plot

We can select the number of TFs to include in the analysis by using `n_tfs` variable.

```{r}
#| fig-width: 16
#| fig-height: 7
#| fig-cap: "Increase the number of reported TFs."

# Increase number of TFs included in the analysis.
out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 n_tfs = 40)
p <- out$heatmaps$average_scores
p
```

## Feature plots of the scores

Perhaps we are interested into visualizing the scores as a Feature plot. This way we can observe trends of enrichment in the activities. This can be achieved by providing `plot_FeaturePlots = TRUE`.

```{r}
#| fig-width: 14
#| fig-height: 8
#| fig-cap: "Retrieve the feature plot output."

# Retrieve feature plots.
out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 plot_FeaturePlots = TRUE)
p1 <- SCpubr::do_DimPlot(sample)
p2 <- out$feature_plots$GLI2
p <- p1 | p2
p
```

## Geyser plots of the scores

In the same fashion as with the feature plots, we can also visualize the scores as Geyser plots, to get a sense of the distribution of the scores alongside the groups. This can be achieved by providing `plot_GeyserPlots = TRUE`.

```{r}
#| fig-width: 14
#| fig-height: 8
#| fig-cap: "Retrieve the geyser plot output."

# Retrieve Geyser plots.
out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 plot_GeyserPlots = TRUE)
p1 <- SCpubr::do_DimPlot(sample)
p2 <- out$geyser_plots$GLI2
p <- p1 | p2
p
```

## Use non-symmetrical color scales

If one wants to just plot a continuous color scale for the Feature plots and the Geyser plots, this can be achieved by using `enforce_symmetry = FALSE`.

```{r}
#| fig-width: 14
#| fig-height: 16
#| fig-cap: "Apply a non-symmetrical color scale."

# Use non-symmetrical color scale.
out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 plot_GeyserPlots = TRUE,
                                 plot_FeaturePlots = TRUE,
                                 enforce_symmetry = FALSE)
p1 <- out$feature_plots$GLI2
p2 <- out$geyser_plots$GLI2

out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 plot_GeyserPlots = TRUE,
                                 plot_FeaturePlots = TRUE,
                                 enforce_symmetry = TRUE)
p3 <- out$feature_plots$GLI2
p4 <- out$geyser_plots$GLI2

p <- (p1 | p2) / (p3 | p4)
p
```

## Order geysers by the mean

We can also decide not to order the Geyser plots by the mean of the values. We can do that by providing `geyser_order_by_mean = FALSE`.

```{r}
#| fig-width: 14
#| fig-height: 8
#| fig-cap: "Order by mean values."

# Not order Geyser plot by mean values.
out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 plot_GeyserPlots = TRUE,
                                 enforce_symmetry = TRUE,
                                 geyser_order_by_mean = FALSE)
p1 <- out$geyser_plots$GLI2

out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 plot_GeyserPlots = TRUE,
                                 enforce_symmetry = TRUE,
                                 geyser_order_by_mean = TRUE)
p2 <- out$geyser_plots$GLI2

p <- p1 | p2
p
```

## Split the heatmap into different groups

We can also further divide the heatmap into groups. This can be achieved using `split.by` parameter.

```{r}
#| fig-width: 12
#| fig-height: 7
#| fig-cap: "Split the heatmap by another variable."

# Split the heatmap by another variable.
sample$split.me <- ifelse(sample$seurat_clusters %in% c("0", "3", "7"), "Group A","Group B")

out <- SCpubr::do_TFActivityPlot(sample = sample,
                                 activities = activities,
                                 split.by = "split.me")
p <- out$heatmaps$average_scores
p
```

## Control the color scale

Again, we can control the color scale using `min.cutoff` and `max.cutoff`. This will apply to all plot types:

```{r}
#| fig-width: 12
#| fig-height: 8
#| fig-cap: "Control the color scale - heatmap."

# Control the color scale
out <- SCpubr::do_TFActivityPlot(sample = sample,
                                      activities = activities,
                                      split.by = "split.me",
                                      min.cutoff = 0.1,
                                      max.cutoff = 0.7,
                                      plot_FeaturePlots = TRUE,
                                      plot_GeyserPlots = TRUE)
p <- out$heatmaps$average_scores
p
```

\`\`\`{r} #\| fig-width: 14 #\| fig-height: 8 #\| fig-cap: "Control the color scale - FeaturePlot and GeyserPlot"

p1 \<- out$feature_plots$PI3K p2 \<- out$geyser_plots$PI3K p \<- p1 \| p2 p
