# Dataset used {.unnumbered}

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Throughout this manual, we use a [publicly available dataset](https://www.10xgenomics.com/resources/datasets/10k-human-pbmcs-3-ht-v3-1-chromium-x-3-1-high) containing 10,000 human PBMCs (peripheral blood mononuclear cells) from 10x Genomics.

Below is the code to generate a Seurat object ready for plotting, including basic quality control, normalization, dimensional reduction, and clustering:

```{r}
#| eval = FALSE
counts_path <- "path_to_count_matrix"

# Read count matrix
counts <- Seurat::Read10X(counts_path)

# Create Seurat object
sample <- Seurat::CreateSeuratObject(counts = counts, project = "10K_pbmc")

# Compute percentage of mitochondrial RNA
sample <- Seurat::PercentageFeatureSet(sample, pattern = "^MT-", col.name = "percent.mt")

# Basic QC filters
mask1 <- sample$nCount_RNA >= 1000
mask2 <- sample$nFeature_RNA >= 500
mask3 <- sample$percent.mt <= 20
mask <- mask1 & mask2 & mask3
sample <- sample[, mask]

# Normalize
sample <- Seurat::SCTransform(sample)

# Dimensional reduction
sample <- Seurat::RunPCA(sample)
sample <- Seurat::RunUMAP(sample, dims = 1:30)

# Find clusters
sample <- Seurat::FindNeighbors(sample, dims = 1:30)
sample <- Seurat::FindClusters(sample, resolution = 0.2)
```

This object will be used as `sample` in all code examples throughout the book.
