# Shared features across functions

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Most functions in `SCpubr` share common parameters. Understanding these will help you navigate the package more efficiently.

## Basic parameters

-   `sample`: **Seurat object** to use for plotting.
-   `group.by`: **Metadata** variable to **group** the values by.
-   `split.by`: **Metadata** variable to **split** the plots by.
-   `features`: Genes or metadata columns to plot.
-   `assay`: **Assay** name to pull data from.
-   `reduction`: Which **dimensional reduction** to use (e.g., "umap", "pca").

## Customization parameters

-   `colors.use`: Custom color palette (vector of named colors).
-   `legend.position`: Where to place the legend ("right", "bottom", "none").
-   `plot.title`: Title for the plot.
-   `plot.subtitle`: Subtitle for the plot.
-   `font.size`: Base font size for the plot.

## Using custom colors

```{r}
#| eval: false
# Define custom colors
colors <- c("0" = "#001219",
            "1" = "#005f73",
            "2" = "#0a9396")

# Apply to any function
p <- SCpubr::do_DimPlot(sample = sample,
                        colors.use = colors)
```

## Colorblind-safe palettes

Set `colorblind = TRUE` in any function to use colorblind-safe palettes automatically:

```{r}
#| eval: false
p <- SCpubr::do_DimPlot(sample = sample,
                        colorblind = TRUE)
```
