# do_ColorPalette() | Color palette generation

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

This utility function generates **harmonious color palettes** based on color theory principles. Provide a base color and get complementary, triadic, tetradic, or other color schemes.

---

## Color wheel

Generate n colors evenly spaced around the color wheel:

```{r}
#| eval: true
colors <- SCpubr::do_ColorPalette(colors.use = "#E63946",
                                  n = 8)
colors
```

---

## Opposite (complementary)

```{r}
#| eval: true
colors <- SCpubr::do_ColorPalette(colors.use = "#E63946",
                                  opposite = TRUE)
colors
```

---

## Adjacent colors

```{r}
#| eval: true
colors <- SCpubr::do_ColorPalette(colors.use = "#E63946",
                                  adjacent = TRUE)
colors
```

---

## Triadic scheme

```{r}
#| eval: true
colors <- SCpubr::do_ColorPalette(colors.use = "#E63946",
                                  triadic = TRUE)
colors
```

---

## Tetradic scheme

```{r}
#| eval: true
colors <- SCpubr::do_ColorPalette(colors.use = "#E63946",
                                  tetradic = TRUE)
colors
```

---

## Square scheme

```{r}
#| eval: true
colors <- SCpubr::do_ColorPalette(colors.use = "#E63946",
                                  square = TRUE)
colors
```

---

## Split complementary

```{r}
#| eval: true
colors <- SCpubr::do_ColorPalette(colors.use = "#E63946",
                                  split_complementary = TRUE)
colors
```


---

## Generate plot

```{r}
#| eval: true
p <- SCpubr::do_ColorPalette(colors.use = "#E63946",
                             triadic = TRUE,
                             plot = TRUE)
p
```

---

## Parameter reference

| Parameter | Description | Default |
|-----------|-------------|---------|
| `colors.use` | Base color (name or HEX) | — |
| `n` | Colors in wheel | `12` |
| `opposite` | Complementary color | `FALSE` |
| `adjacent` | Adjacent colors | `FALSE` |
| `triadic` | Triadic scheme | `FALSE` |
| `split_complementary` | Split complementary | `FALSE` |
| `tetradic` | Tetradic scheme | `FALSE` |
| `square` | Square scheme | `FALSE` |
| `plot` | Return plot | `FALSE` |

---

## See also

- [do_ColorBlindCheck()](02_colorblindcheck.qmd) — Test accessibility
- [Shared features](../00_getting_started/04_shared_features.qmd) — Color scales
