# do_ColorBlindCheck() | Colorblind accessibility testing

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

This function simulates **colorblind vision** to verify your color palette is accessible. Shows how your colors appear under different types of color vision deficiency.

---

## Basic usage

```{r}
#| eval: true
colors <- c("#E63946", "#F1FAEE", "#A8DADC", "#457B9D", "#1D3557")

p <- SCpubr::do_ColorBlindCheck(colors.use = colors)
p
```

---

## Simulated conditions

The function displays your colors as seen by:

- **Normal** — Typical color vision
- **Protanopia** — Red-green (less common)
- **Deuteranopia** — Red-green (most common)
- **Tritanopia** — Blue-yellow

---

## Flip orientation

```{r}
#| eval: true
p <- SCpubr::do_ColorBlindCheck(colors.use = colors,
                                flip = TRUE)
p
```

---

## Parameter reference

| Parameter | Description | Default |
|-----------|-------------|---------|
| `colors.use` | Colors to check | — |
| `flip` | Flip orientation | `FALSE` |

---

## See also

- [do_ColorPalette()](01_colorpalette.qmd) — Generate palettes
- [Shared features](../00_getting_started/04_shared_features.qmd) — Color scales
