# do_MetadataHeatmap() | Metadata overview

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Metadata heatmaps display **categorical metadata variables** across samples or groups. Ideal for visualizing patient characteristics, experimental conditions, or sample annotations.

---

## Basic usage

```{r}
#| eval: true
p <- SCpubr::do_MetadataHeatmap(sample = sample,
                                group.by = "ID",
                                metadata = c("technology", "gender", "subtype"),
                                flip = FALSE,
                                legend.ncol = 1)
p
```

---

## Cluster samples

```{r}
#| eval: true
p <- SCpubr::do_MetadataHeatmap(sample = sample,
                                group.by = "ID",
                                metadata = c("technology", "gender", "subtype"),
                                legend.ncol = 1,
                                flip = FALSE,
                                cluster = TRUE)
p
```

---

## From data frame

```{r}
#| eval: true
df <- data.frame(
  Patient1 = c("Healthy", "Male"),
  Patient2 = c("Disease", "Female"),
  row.names = c("Condition", "Sex")
)

p <- SCpubr::do_MetadataHeatmap(from_df = TRUE,
                                df = df)
p
```

---

## Adjust gap between rows

```{r}
#| eval: true
p <- SCpubr::do_MetadataHeatmap(sample = sample,
                                group.by = "ID",
                                metadata = c("technology", "gender", "subtype"),
                                flip = FALSE,
                                legend.ncol = 1,
                                heatmap.gap = 5)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `group.by` | Column defining groups | — |
| `metadata` | Metadata columns to show | — |
| `colors.use` | Named list of color mappings | `NULL` |

### Layout

| Parameter | Description | Default |
|-----------|-------------|---------|
| `cluster` | Hierarchical clustering | `FALSE` |
| `flip` | Vertical layout | `TRUE` |
| `heatmap.gap` | Gap between rows (mm) | `1` |

### From data frame

| Parameter | Description | Default |
|-----------|-------------|---------|
| `from_df` | Use data frame input | `FALSE` |
| `df` | Data frame with metadata | `NULL` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_ExpressionHeatmap()](../02_gene_expression/05_expressionheatmap.qmd) — Expression heatmaps
- [do_BarPlot()](../03_proportions/01_barplot.qmd) — Proportional bars
