# do_FeaturePlot() | Continuous values on dimensional reduction

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Feature plots display **continuous values**—gene expression, metadata scores, or dimensional reduction components—on a dimensional reduction embedding. Unlike `do_DimPlot()` which colors by categories, `do_FeaturePlot()` maps numeric values to a color gradient.


---

## Basic usage

Plot a single feature (gene, metadata column, or PC component):

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1")
p
```

### Multiple features

Plot several features at once:

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = c("PC_1", "PC_2"),
                            ncol = 2)
p
```

---

## Scale customization

### Control color scale limits

Outliers often compress the color range. Use `min.cutoff` and `max.cutoff` to clamp values:

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            min.cutoff = -10,
                            max.cutoff = 10)
p
```

::: {.callout-tip}
Provide one value per feature: `min.cutoff = c(0, 0.5, 1)` for three features.

:::

### Fixed scale limits

Use `scale.limits` to extend the scale beyond data range (useful for comparing across samples):

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            scale.limits = c(-50, 50))
p
```

::: {.callout-note}
`scale.limits` cannot be used with `min.cutoff` or `max.cutoff`.
:::

### Symmetrical scales

Center the color scale around zero (or another value) for diverging data:

```{r}
#| eval: true
# Absolute symmetry: scale extends equally in both directions
p1 <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            enforce_symmetry = TRUE,
                            symmetry.type = "absolute")

# Centered symmetry: center around a specific value
p2 <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            enforce_symmetry = TRUE,
                            symmetry.type = "centered",
                            symmetry.center = 10)

p1 | p2
```

---

## Color palettes

### Sequential palettes (default)

```{r}
#| eval: true
# RColorBrewer sequential palette
p1 <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            use_viridis = FALSE,
                            sequential.palette = "YlOrRd",
                            sequential.direction = 1)

# Viridis palette
p2 <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            use_viridis = TRUE,
                            viridis.palette = "G",
                            viridis.direction = 1)

p1 | p2
```

### Diverging palettes

Automatically used when `enforce_symmetry = TRUE`:

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            enforce_symmetry = TRUE,
                            diverging.palette = "RdBu",
                            diverging.direction = -1)
p
```

---

## Subsetting & splitting

### Highlight specific cells

```{r}
#| eval: true
# Highlight by cell barcodes
cells.use <- sample(colnames(sample), 500)
p1 <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            cells.highlight = cells.use)

# Highlight by identity
p2 <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            idents.highlight = c("IPC-like"))

p1 | p2
```

### Split by variable

Create faceted plots:

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            split.by = "subtype")
p
```

---

## Group overlay

Display categorical group information alongside continuous values:

### Group dots

Add dots at cluster centers:

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            group.by = "Annotation")
p
```

### Group cell borders

Add a colored shadow around cells showing group membership:

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            group.by = "Annotation",
                            group.by.cell_borders = TRUE,
                            group.by.cell_borders.alpha = 0.1)
p
```

---

## Labels

Add identity labels to the plot:

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            label = TRUE,
                            label.color = "black",
                            label.size = 4)
p
```

---

## Visual enhancements


### Density contours

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "PC_1",
                            plot_density_contour = TRUE,
                            contour.color = "grey50",
                            contour.position = "bottom")
p
```

### Order by expression

Plot cells with higher expression on top:

```{r}
#| eval: true
p <- SCpubr::do_FeaturePlot(sample = sample,
                            features = "KIF14",
                            order = TRUE)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, titles, cell borders, density contours, rendering), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `features` | Features to plot (genes, metadata, PCs) | — |

### Scale & limits (FeaturePlot-specific)

| Parameter | Description | Default |
|-----------|-------------|---------|
| `scale.limits` | Fixed scale limits `c(min, max)` | `NULL` |
| `symmetry.type` | `"absolute"` or `"centered"` | `"absolute"` |
| `symmetry.center` | Center value for `"centered"` type | `NA` |

### Subsetting

| Parameter | Description | Default |
|-----------|-------------|---------|
| `idents.keep` | Identities to display | `NULL` |
| `cells.highlight` | Cell barcodes to highlight | `NULL` |
| `idents.highlight` | Identities to highlight | `NULL` |

### Group overlay

| Parameter | Description | Default |
|-----------|-------------|---------|
| `group.by` | Metadata column for group overlay | `NULL` |
| `group.by.colors.use` | Colors for groups | Auto-generated |
| `group.by.legend` | Legend title for groups | Column name |
| `group.by.dot.size` | Size of center dots | `8` |
| `group.by.show.dots` | Show center dots | `TRUE` |
| `group.by.cell_borders` | Show colored borders | `FALSE` |
| `group.by.cell_borders.alpha` | Border transparency | `0.1` |
| `colorblind` | Colorblind-safe group colors | `FALSE` |

### Individual customization

| Parameter | Description | Default |
|-----------|-------------|---------|
| `individual.titles` | Title per feature | `NULL` |
| `individual.subtitles` | Subtitle per feature | `NULL` |
| `individual.captions` | Caption per feature | `NULL` |

### Appearance (FeaturePlot-specific)

| Parameter | Description | Default |
|-----------|-------------|---------|
| `order` | Order cells by expression | `FALSE` |
| `ncol` | Columns for multiple features | `NULL` |
| `plot.axes` | Show axis lines | `FALSE` |
| `verbose` | Show messages | `TRUE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters across all functions
- [do_DimPlot()](../01_dimensionality_reduction/01_dimplot.qmd) — Categorical visualization on dimensional reductions
- [do_NebulosaPlot()](02_nebulosaplot.qmd) — Density-based feature visualization
