# do_CellularStatesPlot() | Cellular state enrichment plots

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Cellular states plots visualize the relationship between **enrichment scores** for gene signatures. Cells are positioned based on their enrichment in 2, 3, or 4 gene lists, revealing cellular state transitions and plasticity.

This plot is inspired by the following publications:

- **Neftel et al. (2019)** — [Cell 178, 835-849](https://doi.org/10.1016/j.cell.2019.06.024) — Glioblastoma cellular states
- **Tirosh et al. (2016)** — [Nature 539, 309-313](https://doi.org/10.1038/nature20123) — Oligodendroglioma hierarchy


---

## Basic usage

### Two-variable plot

The simplest case: compare enrichment between two signatures.

```{r}
#| eval: true
# Define gene signatures
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157")
)

p <- SCpubr::do_CellularStatesPlot(
  sample          = sample,
  input_gene_list = gene_lists,
  x1              = "IPC.like",
  y1              = "Cilia.like"
)
p
```

Cells are positioned by their enrichment scores: X-axis shows one signature, Y-axis shows the other.

---

## Three-variable plot

Add a second X-axis variable to create a **butterfly plot**. The X-axis becomes a dual axis where:

- **Right side** → enriched in `x1`
- **Left side** → enriched in `x2`  
- **Y-axis** → enrichment in `y1`

```{r}
#| eval: true
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "DNER3", "SNTG1")
)

p <- SCpubr::do_CellularStatesPlot(
  sample          = sample,
  input_gene_list = gene_lists,
  x1              = "Cilia.like",
  x2              = "OPC.like",
  y1              = "IPC.like"
)
p
```

---

## Four-variable plot (quadrant plot)

With four signatures, each corner represents one cellular state:

- **Top-right** → `x1`
- **Top-left** → `x2`
- **Bottom-right** → `y1`
- **Bottom-left** → `y2`

Cells cluster toward corners they're enriched in, or fall between corners if enriched in multiple states.

```{r}
#| eval: true
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "DNER3", "SNTG1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ")
)

p <- SCpubr::do_CellularStatesPlot(
  sample          = sample,
  input_gene_list = gene_lists,
  x1              = "IPC.like",
  x2              = "Mesenchymal.like",
  y1              = "OPC.like",
  y2              = "Cilia.like"
)
p
```

---

## Enforce symmetry

Force the plot to be symmetrical around zero:

```{r}
#| eval: true
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "DNER3", "SNTG1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ")
)

p <- SCpubr::do_CellularStatesPlot(
  sample          = sample,
  input_gene_list = gene_lists,
  x1              = "IPC.like",
  x2              = "Mesenchymal.like",
  y1              = "OPC.like",
  y2              = "Cilia.like",
  enforce_symmetry = TRUE
)
p
```

For 3-variable plots, this centers the X-axis at 0. For 4-variable plots, this creates a square plot with equal ranges on both axes.

---

## Color by metadata

Color cells by any metadata variable:

```{r}
#| eval: true
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "DNER3", "SNTG1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ")
)

p <- SCpubr::do_CellularStatesPlot(
  sample          = sample,
  input_gene_list = gene_lists,
  x1              = "IPC.like",
  x2              = "Mesenchymal.like",
  y1              = "OPC.like",
  y2              = "Cilia.like",
  group.by =  "subtype",
)
p
```

---

## Plot enrichment scores

Visualize the individual enrichment scores as additional feature plots:
 
```{r}
#| eval: true
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "DNER3", "SNTG1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ")
)

p <- SCpubr::do_CellularStatesPlot(
  sample          = sample,
  input_gene_list = gene_lists,
  x1              = "IPC.like",
  x2              = "Mesenchymal.like",
  y1              = "OPC.like",
  y2              = "Cilia.like",
  plot_enrichment_scores = TRUE
)
p$main | ((p$IPC.like | p$Mesenchymal.like) / (p$Cilia.like | p$OPC.like))

```

---

## Overlay additional features

Plot any feature (genes, metadata) onto the cellular states coordinates:

```{r}
#| eval: true
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "DNER3", "SNTG1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ")
)

p <- SCpubr::do_CellularStatesPlot(
  sample          = sample,
  input_gene_list = gene_lists,
  x1              = "IPC.like",
  x2              = "Mesenchymal.like",
  y1              = "OPC.like",
  y2              = "Cilia.like",
  plot_features = TRUE,
  features = c("PC_1"),
)

p$main | p$PC_1
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, titles, cell borders, marginal distributions, rendering), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `input_gene_list` | Named list of gene signatures | — |
| `x1` | First X-axis signature name | — |
| `y1` | Y-axis signature name (or bottom-right quadrant) | — |
| `x2` | Second X-axis signature (enables 3-variable mode) | `NULL` |
| `y2` | Second Y-axis signature (enables 4-variable mode) | `NULL` |

### Enrichment computation

| Parameter | Description | Default |
|-----------|-------------|---------|
| `nbin` | Number of bins for enrichment scoring | `24` |
| `ctrl` | Number of control genes per bin | `100` |

### Plot modes

| Parameter | Description | Default |
|-----------|-------------|---------|
| `enforce_symmetry` | Force symmetrical axis ranges | `FALSE` |
| `plot_enrichment_scores` | Return enrichment score plots | `FALSE` |
| `plot_features` | Return additional feature plots | `FALSE` |
| `features` | Features to plot (when `plot_features = TRUE`) | `NULL` |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `axis.ticks` | Show axis ticks | `TRUE` |
| `axis.text` | Show axis text | `TRUE` |
| `colorblind` | Use colorblind-safe palette | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_FeaturePlot()](01_featureplot.qmd) — Standard gene expression visualization
- [do_EnrichmentHeatmap()](06_enrichmentheatmap.qmd) — Enrichment score heatmaps
