# do_DotPlot() | Marker gene expression

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Dot plots visualize gene expression using **two visual scales**: color represents average expression, while dot size shows the percentage of cells expressing each gene. This dual encoding makes them ideal for comparing marker genes across clusters or conditions.

---

## Basic usage

Plot marker genes across clusters:

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = c("CDC25C", "KIF18B", "KIF14", "CENPE"))
p
```

### Query multiple markers

```{r}
#| eval: true
genes <- c("CDC25C", "KIF18B", "KIF14", "CENPE",
           "DNAAF1", "ADGB", "CFAP61", "CFAP157",
           "KCNQ5", "MEOX2", "MBP", "SNTG1",
           "S100A1", "MGP", "TNNT1", "H2AFJ")

p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes)
p
```

---

## Grouped features

Provide a **named list** to organize genes into categories with faceted panels:

```{r}
#| eval: true
genes <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "MBP", "SNTG1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ")
)

p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes)
p
```

---

## Clustering

### Cluster identities

Group similar clusters together based on expression patterns:

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes,
                        cluster.identities = TRUE)
p
```

### Cluster features

Reorder genes by expression similarity:

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes,
                        cluster.features = TRUE)
p
```

### Cluster both

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes,
                        cluster.identities = TRUE,
                        cluster.features = TRUE)
p
```

---

## Z-score normalization

Show relative expression across groups instead of absolute values. This highlights which group has the highest expression for each gene, but prevents cross-gene comparisons:

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes,
                        zscore.data = TRUE)
p
```

::: {.callout-note}
Z-scoring automatically enables `enforce_symmetry = TRUE` for a centered diverging color scale.
:::

---

## Split by condition

Create separate facets for each condition:

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
                        split.by = "subtype")
p
```

::: {.callout-warning}
`split.by` cannot be combined with named feature lists, clustering, or z-scoring.
:::

---

## Filter low expression

Remove dots below a minimum percent expressed threshold:

```{r}
#| eval: true
# Only show dots where >= 50% of cells express the gene
p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes,
                        dot.min = 50)
p
```

---

## Visual customization

### Flip axes

Swap genes and clusters:

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
                        flip = TRUE)
p
```

### Adjust dot size

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes,
                        dot.scale = 12)
p
```

### Control axis text angle

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes,
                        axis.text.x.angle = 90)
p
```

---

## Scale limits

### Clamp expression values

```{r}
#| eval: true
p <- SCpubr::do_DotPlot(sample = sample,
                        features = genes,
                        min.cutoff = 0,
                        max.cutoff = 3)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, titles, grid, rendering), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `features` | Genes to plot (vector or named list) | — |

### Clustering & transformation

| Parameter | Description | Default |
|-----------|-------------|---------|
| `cluster.identities` | Cluster groups by expression similarity | `FALSE` |
| `cluster.features` | Cluster genes by expression patterns | `FALSE` |
| `zscore.data` | Show Z-scores instead of average expression | `FALSE` |

### Filtering

| Parameter | Description | Default |
|-----------|-------------|---------|
| `dot.min` | Minimum percent expressed to show dot (0–100) | `5` |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `dot.scale` | Maximum dot size | `8` |
| `axis.text.x.angle` | X-axis label rotation angle | `45` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_ExpressionHeatmap()](05_expressionheatmap.qmd) — Heatmap visualization of expression
- [do_ViolinPlot()](../04_distributions/02_violinplot.qmd) — Distribution-based expression visualization
