# do_ExpressionHeatmap() | Averaged expression heatmaps

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Expression heatmaps display **averaged gene expression** across groups. Each cell in the heatmap represents the mean expression of a gene within a group (cluster, condition, etc.), providing a compact summary of expression patterns.


---

## Basic usage

Create an expression heatmap for a set of genes:

```{r}
#| eval: true
genes <- c("CDC25C", "KIF18B", "KIF14", "CENPE",
           "DNAAF1", "ADGB", "CFAP61", "CFAP157",
           "KCNQ5", "MEOX2", "MBP", "SNTG1",
           "S100A1", "MGP", "TNNT1", "H2AFJ")

p <- SCpubr::do_ExpressionHeatmap(sample = sample,
                                  features = genes)
p
```

---

## Multiple grouping variables

Group by multiple metadata variables to create stacked heatmaps:

```{r}
#| eval: true
p <- SCpubr::do_ExpressionHeatmap(sample = sample,
                                  features = genes,
                                  group.by = c("Annotation", "subtype"))
p
```

---

## Clustering

### Automatic clustering (default)

By default, both genes and groups are hierarchically clustered:

```{r}
#| eval: true
p <- SCpubr::do_ExpressionHeatmap(sample = sample,
                                  features = genes,
                                  cluster = TRUE)
p
```

### Disable clustering

Preserve the original order:

```{r}
#| eval: true
p <- SCpubr::do_ExpressionHeatmap(sample = sample,
                                  features = genes,
                                  cluster = FALSE)
p
```

---

## Display values

Show expression values as text overlaid on cells:

```{r}
#| eval: true
p <- SCpubr::do_ExpressionHeatmap(sample = sample,
                                  features = genes,
                                  values.show = TRUE,
                                  values.threshold = 1.5,
                                  values.size = 3,
                                  values.round = 1)
p
```

| Parameter | Description |
|-----------|-------------|
| `values.show` | Display numeric values on cells |
| `values.threshold` | Threshold for switching text color (white above, black below) |
| `values.size` | Text size |
| `values.round` | Decimal places to display |

---

## Scale customization

### Clamp values

```{r}
#| eval: true
p <- SCpubr::do_ExpressionHeatmap(sample = sample,
                                  features = genes,
                                  min.cutoff = 0,
                                  max.cutoff = 1.5)
p
```

### Symmetrical (diverging) scale

Center the color scale around zero:

```{r}
#| eval: true
p <- SCpubr::do_ExpressionHeatmap(sample = sample,
                                  features = genes,
                                  enforce_symmetry = TRUE,
                                  diverging.palette = "RdBu")
p
```

---

## Visual customization


### Customize grid and border colors

```{r}
#| eval: true
p <- SCpubr::do_ExpressionHeatmap(sample = sample,
                                  features = genes,
                                  grid.color = "blue",
                                  border.color = "red")
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `features` | Genes to display | — |

### Clustering & ordering

| Parameter | Description | Default |
|-----------|-------------|---------|
| `cluster` | Hierarchically cluster genes and groups | `TRUE` |
| `features.order` | Custom gene order | `NULL` |
| `groups.order` | Custom group order (named list) | `NULL` |

### Value display

| Parameter | Description | Default |
|-----------|-------------|---------|
| `values.show` | Show values on cells | `FALSE` |
| `values.threshold` | Color switch threshold | `NULL` |
| `values.size` | Value text size | `3` |
| `values.round` | Decimal places | `1` |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `grid.color` | Color between cells | `"white"` |
| `border.color` | Heatmap border color | `"black"` |
| `axis.text.x.angle` | X-axis label rotation | `45` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_EnrichmentHeatmap()](06_enrichmentheatmap.qmd) — Enrichment score heatmaps
- [do_SCExpressionHeatmap()](07_scexpressionheatmap.qmd) — Single-cell level expression heatmaps
