# do_EnrichmentHeatmap() | Gene set enrichment heatmaps

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Enrichment heatmaps display **averaged enrichment scores** for gene sets across groups. This function computes enrichment scores using Seurat's `AddModuleScore` (or UCell) and aggregates them by metadata variables.

---

## Basic usage

```{r}
#| eval: true
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "MBP", "SNTG1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ")
)

p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = gene_lists,
                                  flip = TRUE)
p
```

---

## Multiple grouping variables

```{r}
#| eval: true
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = gene_lists,
                                  group.by = c("Annotation", "subtype"),
                                  flip = TRUE)
p
```

---

## Scoring methods
Use `flavor = "Seurat"` or `flavor = "UCell"` to toggle different scoring methods.


---

## Scale scores

Transform scores to 0-1 range for easier comparison:

```{r}
#| eval: true
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = gene_lists,
                                  scale_scores = TRUE, 
                                  flip = TRUE)
p
```

---

## Display values

```{r}
#| eval: true
p <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                  input_gene_list = gene_lists,
                                  values.show = TRUE,
                                  values.threshold = 0.5,
                                  flip = TRUE)
p
```

---

## Return enriched object

Get the Seurat object with computed enrichment scores:

```{r}
#| eval: true
result <- SCpubr::do_EnrichmentHeatmap(sample = sample,
                                       input_gene_list = gene_lists,
                                       return_object = TRUE)

p <- result$plot
sample_enriched <- result$object
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `input_gene_list` | Named list of gene signatures | — |

### Scoring method

| Parameter | Description | Default |
|-----------|-------------|---------|
| `flavor` | Scoring method (`"Seurat"` or `"UCell"`) | `"Seurat"` |
| `nbin` | Number of bins (Seurat) | `24` |
| `ctrl` | Control genes per bin (Seurat) | `100` |
| `ncores` | CPU cores for UCell | `1` |
| `storeRanks` | Store ranks for faster UCell | `TRUE` |
| `scale_scores` | Scale scores to 0-1 range | `FALSE` |

### Clustering & ordering

| Parameter | Description | Default |
|-----------|-------------|---------|
| `cluster` | Hierarchically cluster | `TRUE` |
| `features.order` | Custom gene set order | `NULL` |
| `groups.order` | Custom group order | `NULL` |

### Output

| Parameter | Description | Default |
|-----------|-------------|---------|
| `return_object` | Return Seurat object with scores | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_ExpressionHeatmap()](05_expressionheatmap.qmd) — Gene expression heatmaps
- [do_SCEnrichmentHeatmap()](08_scenrichmentheatmap.qmd) — Single-cell enrichment heatmaps
