# do_SCExpressionHeatmap() | Single-cell expression heatmaps

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Single-cell expression heatmaps display gene expression at **individual cell resolution**. Cells are ordered by group, and optional metadata tracks can be added alongside the main heatmap. Inspired by Seurat's `DoHeatmap`.

---

## Basic usage

```{r}
#| eval: true
genes <- c("CDC25C", "KIF18B", "KIF14", "CENPE",
           "DNAAF1", "ADGB", "CFAP61", "CFAP157",
           "KCNQ5", "MEOX2", "MBP", "SNTG1",
           "S100A1", "MGP", "TNNT1", "H2AFJ")

p <- SCpubr::do_SCExpressionHeatmap(sample = sample,
                                    features = genes)
p
```

---

## Add metadata tracks

Display categorical metadata alongside the expression heatmap:

```{r}
#| eval: true
p <- SCpubr::do_SCExpressionHeatmap(
  sample          = sample,
  features        = genes,
  metadata        = c("Annotation", "subtype")
)
p
```

---

## Clustering

### Cluster genes

```{r}
#| eval: true
p <- SCpubr::do_SCExpressionHeatmap(sample = sample,
                                    features = genes,
                                    cluster = TRUE)
p
```


---

## Subsampling

For large datasets, subsample cells to reduce memory and improve rendering:

```{r}
#| eval: true
p <- SCpubr::do_SCExpressionHeatmap(sample = sample,
                                    features = genes,
                                    subsample = 5)
p
```

---

## Proportional sizing

### Equal space per group (default)

```{r}
#| eval: true
p <- SCpubr::do_SCExpressionHeatmap(sample = sample,
                                    features = genes,
                                    proportional.size = TRUE)
p
```

### Size proportional to cell count

```{r}
#| eval: true
p <- SCpubr::do_SCExpressionHeatmap(sample = sample,
                                    features = genes,
                                    proportional.size = FALSE)
p
```

---

## Interpolation

Smooth the heatmap for rasterized output:

```{r}
#| eval: true
p <- SCpubr::do_SCExpressionHeatmap(sample = sample,
                                    features = genes,
                                    interpolate = TRUE)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `features` | Genes to display | — |

### Metadata tracks

| Parameter | Description | Default |
|-----------|-------------|---------|
| `metadata` | Metadata columns to display | `NULL` |
| `metadata.colors` | Named list of colors for metadata | `NULL` |

### Layout

| Parameter | Description | Default |
|-----------|-------------|---------|
| `proportional.size` | Equal space per group | `TRUE` |
| `main.heatmap.size` | Main heatmap proportion (0-1) | `0.95` |
| `strip.spacing` | Space between group strips | `10` |
| `strip.text.angle` | Strip label rotation | `0` |

### Subsampling & clustering

| Parameter | Description | Default |
|-----------|-------------|---------|
| `subsample` | Max cells per group | `NA` |
| `cluster` | Cluster genes | `TRUE` |
| `features.order` | Custom gene order | `NULL` |
| `interpolate` | Smooth heatmap | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_ExpressionHeatmap()](05_expressionheatmap.qmd) — Averaged expression heatmaps
- [do_SCEnrichmentHeatmap()](08_scenrichmentheatmap.qmd) — Single-cell enrichment heatmaps
