# do_SCEnrichmentHeatmap() | Single-cell enrichment heatmaps

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Single-cell enrichment heatmaps display **enrichment scores at individual cell resolution**. Similar to `do_SCExpressionHeatmap` but for gene set scores instead of individual genes.

---

## Basic usage

```{r}
#| eval: true
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "MBP", "SNTG1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ")
)

p <- SCpubr::do_SCEnrichmentHeatmap(sample = sample,
                                    input_gene_list = gene_lists)
p
```

---

## Add metadata tracks

```{r}
#| eval: true
p <- SCpubr::do_SCEnrichmentHeatmap(
  sample          = sample,
  input_gene_list = gene_lists,
  metadata        = c("Annotation", "subtype")
)
p
```

---

## Scoring methods

Use `flavor = "Seurat"` or `flavor = "UCell"` to toggle different scoring methods.

---

## Subsampling

```{r}
#| eval: true
p <- SCpubr::do_SCEnrichmentHeatmap(sample = sample,
                                    input_gene_list = gene_lists,
                                    subsample = 5)
p
```

---

## Return enriched object

```{r}
#| eval: true
result <- SCpubr::do_SCEnrichmentHeatmap(sample = sample,
                                         input_gene_list = gene_lists,
                                         return_object = TRUE)

p <- result$plot
sample_enriched <- result$object
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `input_gene_list` | Named list of gene signatures | — |

### Scoring method

| Parameter | Description | Default |
|-----------|-------------|---------|
| `flavor` | `"Seurat"` or `"UCell"` | `"Seurat"` |
| `nbin` | Bins for Seurat scoring | `24` |
| `ctrl` | Control genes per bin | `100` |
| `ncores` | CPU cores for UCell | `1` |

### Metadata tracks

| Parameter | Description | Default |
|-----------|-------------|---------|
| `metadata` | Metadata columns to display | `NULL` |
| `metadata.colors` | Named list of colors | `NULL` |

### Layout

| Parameter | Description | Default |
|-----------|-------------|---------|
| `proportional.size` | Equal space per group | `TRUE` |
| `main.heatmap.size` | Main heatmap proportion | `0.95` |
| `subsample` | Max cells per group | `NA` |
| `cluster` | Cluster gene sets | `TRUE` |
| `return_object` | Return Seurat object | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_EnrichmentHeatmap()](06_enrichmentheatmap.qmd) — Averaged enrichment heatmaps
- [do_SCExpressionHeatmap()](07_scexpressionheatmap.qmd) — Single-cell expression heatmaps
