# do_RankedExpressionHeatmap() | Expression along components

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Ranked expression heatmaps display gene expression **ordered by a dimensional reduction component**. Cells are ranked along a PC, UMAP dimension, or trajectory score, revealing expression patterns along continuous gradients.

---

## Basic usage

Rank cells by the first PC component:

```{r}
#| eval: true
genes <- c("CDC25C", "KIF18B", "KIF14", "CENPE",
           "DNAAF1", "ADGB", "CFAP61", "CFAP157",
           "KCNQ5", "MEOX2", "MBP", "SNTG1",
           "S100A1", "MGP", "TNNT1", "H2AFJ")

p <- SCpubr::do_RankedExpressionHeatmap(sample = sample,
                                        features = genes,
                                        reduction = "pca",
                                        dims = 1)
p
```

---

## Multiple dimensions

Generate one heatmap per dimension:

```{r}
#| eval: true
p <- SCpubr::do_RankedExpressionHeatmap(sample = sample,
                                        features = genes,
                                        reduction = "pca",
                                        dims = 1:2)

# Access individual plots
p$PC_1 | p$PC_2
```

---

## Add metadata track

Overlay group information:

```{r}
#| eval: true
p <- SCpubr::do_RankedExpressionHeatmap(
  sample     = sample,
  features   = genes,
  reduction = "pca",
  dims       = 1,
  group.by   = c("Annotation", "subtype"),
)
p
```

---

## Subsampling

Reduce cells for faster plotting:

```{r}
#| eval: true
p <- SCpubr::do_RankedExpressionHeatmap(sample = sample,
                                        features = genes,
                                        dims = 1,
                                        subsample = 50)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `features` | Genes to display | — |
| `dims` | Dimensions to rank cells by | `1:2` |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `main.heatmap.size` | Main heatmap proportion (0-1) | `0.95` |
| `subsample` | Max cells to display | `2500` |
| `interpolate` | Smooth heatmap | `FALSE` |

### Metadata

| Parameter | Description | Default |
|-----------|-------------|---------|
| `colors.use` | Named list of group colors | `NULL` |
| `return_object` | Return Seurat object | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_RankedEnrichmentHeatmap()](10_rankedenrichmentheatmap.qmd) — Enrichment along components
- [do_SCExpressionHeatmap()](07_scexpressionheatmap.qmd) — Single-cell expression heatmaps
