# do_RankedEnrichmentHeatmap() | Enrichment along components

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Ranked enrichment heatmaps display **gene set enrichment scores ordered by a dimensional reduction component**. Cells are ranked along a PC or trajectory, revealing how enrichment scores vary along continuous gradients.

---

## Basic usage

```{r}
#| eval: true
gene_lists <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157"),
  "OPC.like" = c("KCNQ5", "MEOX2", "MBP", "SNTG1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ")
)

p <- SCpubr::do_RankedEnrichmentHeatmap(sample = sample,
                                        input_gene_list = gene_lists,
                                        reduction = "pca",
                                        dims = 1)
p
```

---

## Multiple dimensions

```{r}
#| eval: true
p <- SCpubr::do_RankedEnrichmentHeatmap(sample = sample,
                                        input_gene_list = gene_lists,
                                        reduction = "pca",
                                        dims = 1:2)

# Access individual plots
p$PC_1 | p$PC_2
```

---

## Scale enrichment scores (default)

Z-score enrichment for easier comparison between gene sets:

```{r}
#| eval: true
p <- SCpubr::do_RankedEnrichmentHeatmap(sample = sample,
                                        input_gene_list = gene_lists,
                                        dims = 1,
                                        reduction = "pca",
                                        scale.enrichment = TRUE)
p
```

::: {.callout-note}
Scaling enables intra-gene set comparison but prevents absolute value comparison between gene sets.
:::

---

## Add metadata track

```{r}
#| eval: true
p <- SCpubr::do_RankedEnrichmentHeatmap(
  sample          = sample,
  input_gene_list = gene_lists,
  dims            = 1,
  reduction = "pca",
  group.by        = c("Annotation", "subtype"),
)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `input_gene_list` | Named list of gene signatures | — |
| `dims` | Dimensions to rank cells by | `1:2` |

### Scoring

| Parameter | Description | Default |
|-----------|-------------|---------|
| `flavor` | `"Seurat"` or `"UCell"` | `"Seurat"` |
| `scale.enrichment` | Z-score enrichment values | `TRUE` |
| `nbin` | Bins for Seurat scoring | `24` |
| `ctrl` | Control genes per bin | `100` |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `main.heatmap.size` | Main heatmap proportion (0-1) | `0.95` |
| `subsample` | Max cells to display | `2500` |
| `return_object` | Return Seurat object | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_RankedExpressionHeatmap()](09_rankedexpressionheatmap.qmd) — Expression along components
- [do_EnrichmentHeatmap()](06_enrichmentheatmap.qmd) — Averaged enrichment heatmaps
