# do_DimPlot() | Dimensional reduction plots

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Dimensional reduction plots (`DimPlots`) are the cornerstone of single-cell visualization. They display cells in a reduced-dimension embedding (UMAP, tSNE, PCA) colored by any categorical variable—clusters, cell types, conditions, or any metadata.

---


## Basic usage

Generate a DimPlot with default settings. Cells are colored by the active identity (`Idents(sample)`):

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample)
p
```

By default, axes are hidden for a cleaner look. To show them:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        plot.axes = TRUE)
p
```

---

## Grouping & splitting

### Group by a metadata variable

Color cells by any metadata column using `group.by`:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        group.by = "subtype")
p
```

### Split into facets

Create separate panels for each value of a variable:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        split.by = "subtype",
                        ncol = 3)
p
```
Note: the error in the first plot is currently being reviewed.


A combined overview panel is automatically added. Control this with `split.by.combined`:

```{r}
#| eval: true
# Without the combined panel
p <- SCpubr::do_DimPlot(sample = sample,
                        split.by = "subtype",
                        split.by.combined = FALSE)
p
```

### Combine group.by and split.by

Color by one variable, split by another:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        group.by = "subtype",
                        split.by = "technology")
p
```

### Restrict displayed identities

Show only specific groups while preserving the full UMAP silhouette:

```{r}
#| eval: true
# Only show clusters 0, 2, and 5
p <- SCpubr::do_DimPlot(sample = sample,
                        idents.keep = c("IPC-like", "Cilia-like", "NPC-like", "OPC-like", "Mesenchymal-like"))
p
```

This is preferable to subsetting the Seurat object, as it retains context from all cells.

---

## Labeling clusters

### Basic labels

Add labels directly on clusters:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        label = TRUE)
p
```

### Label customization

| Parameter | Description | Default |
|-----------|-------------|---------|
| `label` | Show labels | `FALSE` |
| `label.box` | Draw box around label | `TRUE` |
| `label.color` | Text color | `"black"` |
| `label.fill` | Box fill color (`NULL` = match cluster color) | `"white"` |
| `label.size` | Text size | `4` |
| `repel` | Repel overlapping labels | `FALSE` |


---
 
## Highlighting cells

### Highlight specific cells

Provide cell names to highlight:

```{r}
#| eval: true
cells.use <- sample(colnames(sample), 1500)

p <- SCpubr::do_DimPlot(sample = sample,
                        cells.highlight = cells.use)
p
```

### Highlight entire identities

Highlight all cells belonging to specific clusters:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        idents.highlight = c("IPC-like", "Cilia-like"))
p
```

::: {.callout-note}
`cells.highlight` and `idents.highlight` cannot be used with `group.by` or `split.by`.
:::

---

### Density contours

Overlay density contour lines:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        plot_density_contour = TRUE,
                        contour.color = "grey90",
                        contour.position = "top")
p
```

### Marginal distributions

Add density/histogram plots along axes:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        plot_marginal_distributions = TRUE,
                        marginal.type = "density",
                        marginal.group = TRUE,
                        plot_cell_borders = FALSE)
p
```

| `marginal.type` values | Description |
|------------------------|-------------|
| `"density"` | Density curves |
| `"histogram"` | Histograms |
| `"boxplot"` | Box plots |
| `"violin"` | Violin plots |
| `"densigram"` | Density + histogram |

::: {.callout-warning}
Marginal distributions cannot be used with `split.by`, `plot_cell_borders`, or `cells.highlight`.
:::

---

## Colors

### Custom color palette

Provide a named vector of colors:

```{r}
#| eval: true
colors <- c("ATRT-TYR" = "#E63946", "ATRT-SHH" = "#457B9D", "ATRT-MYC" = "#1D3557")

p <- SCpubr::do_DimPlot(sample = sample,
                        group.by = "subtype",
                        colors.use = colors)
p
```

### Colorblind-safe palette

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        colorblind = TRUE)
p
```

### NA value color

Control the color for cells set to NA via `na.value`:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        idents.keep = c("IPC-like"),
                        na.value = "grey25")
p
```

---

## Rasterization

For large datasets, rasterize to reduce file size:

```{r}
#| eval: true
p <- SCpubr::do_DimPlot(sample = sample,
                        raster = TRUE,
                        raster.dpi = 2048,
                        pt.size = 8)
p
```

::: {.callout-tip}
If `pt.size < 1` with `raster = TRUE`, cells appear as crosses. Use `pt.size >= 1`.
:::

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, titles, cell borders, density contours, marginal distributions, rendering), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `sample` | Seurat object | — |
| `reduction` | Dimensional reduction to use | Auto-detected |
| `group.by` | Metadata column to color by | Active idents |
| `split.by` | Metadata column to facet by | `NULL` |
| `split.by.combined` | Add combined overview panel | `TRUE` |
| `dims` | Which dimensions to plot | `c(1, 2)` |
| `ncol` | Number of columns for faceted plots | `NULL` |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `pt.size` | Point size | `1` |
| `colors.use` | Named vector of colors | Auto-generated |
| `colorblind` | Use colorblind-safe palette | `FALSE` |
| `shuffle` | Randomize plotting order | `TRUE` |
| `order` | Identities to plot last (on top) | `NULL` |
| `plot.axes` | Show axis lines and labels | `FALSE` |

### Labels

| Parameter | Description | Default |
|-----------|-------------|---------|
| `label` | Show cluster labels | `FALSE` |
| `label.box` | Draw box around labels | `TRUE` |
| `label.color` | Label text color | `"black"` |
| `label.fill` | Label box fill | `"white"` |
| `label.size` | Label text size | `4` |
| `repel` | Repel overlapping labels | `FALSE` |

### Highlighting

| Parameter | Description | Default |
|-----------|-------------|---------|
| `cells.highlight` | Cell barcodes to highlight | `NULL` |
| `idents.highlight` | Identities to highlight | `NULL` |
| `idents.keep` | Identities to display (others → NA) | `NULL` |
| `sizes.highlight` | Point size for highlighted cells | `1` |

### DimPlot-specific legend options

| Parameter | Description | Default |
|-----------|-------------|---------|
| `legend.dot.border` | Add border to legend dots | `TRUE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters across all functions
- [do_FeaturePlot()](../02_gene_expression/01_featureplot.qmd) — Continuous variable visualization on dimensional reductions
- [do_NebulosaPlot()](../02_gene_expression/02_nebulosaplot.qmd) — Density-based feature visualization
