# do_BoxPlot() | Box-and-whisker plots

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Box plots provide a **statistical summary** of distributions: median, quartiles, and outliers. Use them to compare gene expression or other continuous values across groups with optional statistical testing.

---

## Basic usage

```{r}
#| eval: true
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA",
                        flip = TRUE,
                        order = TRUE)
p
```

---

## Split by condition

```{r}
#| eval: true
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nCount_RNA",
                        group.by = "subtype",
                        split.by = "technology",
                        flip = TRUE)
p
```

---

## Statistical testing

Add significance annotations between groups:

```{r}
#| eval: true
p <- SCpubr::do_BoxPlot(sample = sample,,
                        feature = "nFeature_RNA",
                        group.by = "subtype",
                        flip = TRUE,
                        order = TRUE,
                        use_test = TRUE,
                        comparisons = list(c("ATRT-TYR", "ATRT-SHH")),
                        test = "wilcox.test")
p
```

::: {.callout-note}
`use_test` cannot be combined with `split.by`.
:::

### Customize significance display

```{r}
#| eval: true
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nFeature_RNA",
                        group.by = "subtype",
                        flip = TRUE,
                        order = TRUE,
                        use_test = TRUE,
                        comparisons = list(c("ATRT-TYR", "ATRT-SHH")),
                        test = "wilcox.test",
                        map_signif_level = c("***" = 0.0001, 
                                             "**" = 0.001, 
                                             "*" = 0.05))
p
```

---

## Silhouette style

Color borders instead of fill:

```{r}
#| eval: true
p <- SCpubr::do_BoxPlot(sample = sample,
                        feature = "nFeature_RNA",
                        group.by = "subtype",
                        flip = TRUE,
                        use_silhouette = TRUE)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `feature` | Feature to plot | — |

### Statistical testing

| Parameter | Description | Default |
|-----------|-------------|---------|
| `use_test` | Add significance testing | `FALSE` |
| `comparisons` | List of group pairs to compare | `NULL` |
| `test` | Test method | `"wilcox.test"` |
| `map_signif_level` | P-value to symbol mapping | See defaults |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `use_silhouette` | Color borders instead of fill | `FALSE` |
| `boxplot.line.color` | Border color | `"black"` |
| `boxplot.linewidth` | Border width | `0.5` |
| `boxplot.width` | Box width | `NULL` |
| `outlier.color` | Outlier dot color | `"black"` |
| `outlier.alpha` | Outlier transparency | `0.5` |
| `order` | Order groups by median | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_ViolinPlot()](02_violinplot.qmd) — Violin distribution plots
- [do_BeeSwarmPlot()](05_beeswarmplot.qmd) — Individual cell plots
