# do_ViolinPlot() | Distribution comparison

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Violin plots show the **full distribution shape** of gene expression across groups. They combine the benefits of box plots and kernel density plots.


---

## Basic usage

```{r}
#| eval: true
p <- SCpubr::do_ViolinPlot(sample = sample,
                           features = "nFeature_RNA",
                           group.by = "subtype")
p
```

---

## Multiple features

```{r}
#| eval: true
p <- SCpubr::do_ViolinPlot(sample = sample,
                           features = c("nCount_RNA", "nFeature_RNA"),
                           group.by = "subtype",
                           ncol = 2)
p
```

### Share Y-axis limits

Force all features to use the same Y-axis range:

```{r}
#| eval: true
p <- SCpubr::do_ViolinPlot(sample = sample,
                           features = c("nCount_RNA", "nFeature_RNA"),
                           group.by = "subtype",
                           ncol = 2,
                           share.y.lims = TRUE)
p
```

---

## Show/hide boxplot

```{r}
#| eval: true

# Without boxplot
p <- SCpubr::do_ViolinPlot(sample = sample,
                           features = "nCount_RNA",
                           group.by = "subtype",
                           plot_boxplot = FALSE)
p
```

---

## Split by variable

```{r}
#| eval: true
p <- SCpubr::do_ViolinPlot(sample = sample,
                           features = "nCount_RNA",
                           group.by = "subtype",
                           split.by = "technology",
                           plot_boxplot = FALSE,
                           order = FALSE)
p
```

---

## Cut Y-axis

Truncate the Y-axis at a specific value:

```{r}
#| eval: true
p <- SCpubr::do_ViolinPlot(sample = sample,
                           features = "nCount_RNA",
                           group.by = "subtype",
                           y_cut = 25000)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `features` | Features to plot | — |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `plot_boxplot` | Show inner boxplot | `TRUE` |
| `boxplot_width` | Width of inner boxplot | `0.2` |
| `line_width` | Violin border width | `0.5` |
| `y_cut` | Y-axis cutoff per feature | `NA` |

### Layout

| Parameter | Description | Default |
|-----------|-------------|---------|
| `ncol` | Columns for multiple features | `NULL` |
| `share.y.lims` | Same Y-axis for all features | `FALSE` |
| `order` | Order groups by median | `TRUE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_BoxPlot()](01_boxplot.qmd) — Box-and-whisker plots
- [do_RidgePlot()](03_ridgeplot.qmd) — Ridge density plots
