# do_RidgePlot() | Density ridge plots

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Ridge plots display **density distributions stacked vertically**, making it easy to compare distributions across many groups. Built on the `ggridges` package.


---

## Basic usage

```{r}
#| eval: true
p <- SCpubr::do_RidgePlot(sample = sample,
                          feature = "nCount_RNA",
                          group.by = "subtype")
p
```

---

## Split by variable

```{r}
#| eval: true
p <- SCpubr::do_RidgePlot(sample = sample,
                          feature = "nCount_RNA",
                          group.by = "subtype",
                          split.by = "technology")
p
```

---

## Continuous color scale

Color ridges by a continuous gradient instead of categorical colors:

```{r}
#| eval: true
p <- SCpubr::do_RidgePlot(sample = sample,
                          feature = "nCount_RNA",
                          group.by = "subtype",
                          continuous_scale = TRUE)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `feature` | Feature to plot | — |
| `continuous_scale` | Use continuous color scale | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_ViolinPlot()](02_violinplot.qmd) — Violin distribution plots
- [do_BeeSwarmPlot()](05_beeswarmplot.qmd) — Individual cell plots
