# do_StripPlot() | Geyser plots

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Strip plots display **continuous values as jittered points** grouped by categories, with summary statistics (mean and confidence intervals) overlaid. Cells can be colored by categorical or continuous variables.

---

## Basic usage

```{r}
#| eval: true
p <- SCpubr::do_StripPlot(sample = sample,
                          features = "nCount_RNA")
p
```

---

## Color by category

```{r}
#| eval: true
p <- SCpubr::do_StripPlot(sample = sample,
                          features = "nCount_RNA",
                          scale_type = "categorical")
p
```

---

## Color by continuous value

```{r}
#| eval: true
p <- SCpubr::do_StripPlot(sample = sample,
                          features = "nCount_RNA",
                          scale_type = "continuous")
p
```

---

## Adjust jitter

Control the horizontal spread of points:

```{r}
#| eval: true
# More compact
p <- SCpubr::do_StripPlot(sample = sample,
                          features = "nCount_RNA",
                          jitter = 0.2)
p
```

---

## Split by variable

```{r}
#| eval: true
p <- SCpubr::do_StripPlot(sample = sample,
                          features = "nCount_RNA",
                          group.by = "subtype",
                          split.by = "technology")
p
```

---

## Symmetrical scale

Center the color scale around zero:

```{r}
#| eval: true
p <- SCpubr::do_StripPlot(sample = sample,
                          features = "PC_1",
                          enforce_symmetry = TRUE)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `features` | Features to plot | — |
| `scale_type` | `"continuous"` or `"categorical"` | `"continuous"` |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `jitter` | Horizontal jitter amount | `0.45` |
| `order` | Order groups by median | `TRUE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_BeeSwarmPlot()](05_beeswarmplot.qmd) — Similar with beeswarm layout
- [do_ViolinPlot()](02_violinplot.qmd) — Violin distribution plots
