# do_BeeSwarmPlot() | Individual cell distributions

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Beeswarm plots display **individual cells as ranked points**, with beeswarm-style jittering to avoid overlaps. Cells are ranked by a feature (e.g., enrichment score) and colored by categorical or continuous values. Built on the `ggbeeswarm` package.

---

## Basic usage

Rank cells by a feature (typically a score in metadata):

```{r}
#| eval: true
p <- SCpubr::do_BeeSwarmPlot(sample = sample,
                             feature_to_rank = "nCount_RNA")
p
```

---

## Continuous vs categorical coloring

### Continuous (color by same feature)

```{r}
#| eval: true
p <- SCpubr::do_BeeSwarmPlot(sample = sample,
                             feature_to_rank = "nCount_RNA",
                             continuous_feature = TRUE)
p
```

### Categorical (color by group)

```{r}
#| eval: true
p <- SCpubr::do_BeeSwarmPlot(sample = sample,
                             feature_to_rank = "nCount_RNA",
                             continuous_feature = FALSE)
p
```

---

## Order groups by median

```{r}
#| eval: true
p <- SCpubr::do_BeeSwarmPlot(sample = sample,
                             feature_to_rank = "nCount_RNA",
                             order = TRUE)
p
```


---

## Remove axes

```{r}
#| eval: true
p <- SCpubr::do_BeeSwarmPlot(sample = sample,
                             feature_to_rank = "nCount_RNA",
                             remove_x_axis = TRUE,
                             remove_y_axis = TRUE)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, cell borders, rendering), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `feature_to_rank` | Feature to rank cells by | — |
| `continuous_feature` | Is feature continuous? | `FALSE` |

### Layout

| Parameter | Description | Default |
|-----------|-------------|---------|
| `order` | Order groups by median | `FALSE` |
| `remove_x_axis` | Hide X-axis | `FALSE` |
| `remove_y_axis` | Hide Y-axis | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_StripPlot()](04_stripplot.qmd) — Similar geyser-style plots
- [do_ViolinPlot()](02_violinplot.qmd) — Violin distribution plots
