# Installation guide {.unnumbered}

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

To install `SCpubr`, use the following commands:

```{r}
#| eval: false
# From CRAN - Official release:
install.packages("SCpubr")

# From GitHub - Latest stable development version:
if(!requireNamespace("devtools", quietly = TRUE)){
  install.packages("devtools") # If not installed.
}

devtools::install_github("enblacar/SCpubr", ref = "v3.0.0")
```

## Check your dependencies

`SCpubr` relies on several packages to run its functions. You can get a full overview of which dependencies are installed and which functions are available using `SCpubr::do_PackageReport()`:

```{r}
#| eval: true
SCpubr::do_PackageReport(extended = TRUE)
```

This will provide you with:

- The status of your **dependencies** (installed or missing).
- The **functions** that can be run based on your currently installed dependencies.
- Packages that require a **minimum version** will be marked with an exclamation mark if outdated.
