# do_ActivityHeatmap() | Activity scores heatmap

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Activity heatmaps display **gene set activity scores** computed using `decoupleR`. Input a named list of gene sets and the function computes activity scores per cell, then aggregates them by group for visualization.

---

## Basic usage

```{r}
#| eval: true
# Define gene sets
gene_sets <- list(
  "IPC.like"      = c("CDC25C", "KIF18B", "KIF14", "CENPE", "POLQ"),
  "Cilia.like" = c("DNAAF1", "ADGB", "CFAP61", "CFAP157", "CFAP46"),
  "OPC.like" = c("KCNQ5", "MEOX2", "MBP", "SNTG1", "KCNIP1"),
  "Mesenchymal.like" = c("S100A1", "MGP", "TNNT1", "H2AFJ", "KCNIP1")
)

p <- SCpubr::do_ActivityHeatmap(sample = sample,
                                input_gene_list = gene_sets)
p
```

---

## Group by custom variable

```{r}
#| eval: true
p <- SCpubr::do_ActivityHeatmap(sample = sample,
                                input_gene_list = gene_sets,
                                group.by = "subtype")
p
```

---

## Subsample cells

```{r}
#| eval: true
p <- SCpubr::do_ActivityHeatmap(sample = sample,
                                input_gene_list = gene_sets,
                                subsample = 1000)
p
```

---

## Show values

```{r}
#| eval: true
p <- SCpubr::do_ActivityHeatmap(sample = sample,
                                input_gene_list = gene_sets,
                                values.show = TRUE,
                                values.threshold = 4)
p
```

---

## Return data object

```{r}
#| eval: false
result <- SCpubr::do_ActivityHeatmap(sample = sample,
                                     input_gene_list = gene_sets,
                                     return_object = TRUE)

# Access the plot
result$plot

# Access activity data
result$activities
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `input_gene_list` | Named list of gene sets | — |
| `statistic` | DecoupleR statistic | `"ulm"` |
| `flavor` | Scoring method: `"Seurat"` or `"UCell"` | `"Seurat"` |
| `return_object` | Return data along with plot | `FALSE` |

### Subsampling

| Parameter | Description | Default |
|-----------|-------------|---------|
| `subsample` | Number of cells to sample | `2500` |
| `nbin` | Number of bins for scoring | `24` |
| `ctrl` | Control genes per bin | `100` |

### Values display

| Parameter | Description | Default |
|-----------|-------------|---------|
| `values.show` | Show numeric values | `FALSE` |
| `values.threshold` | Threshold on which the text changes color | `NULL` |
| `values.round` | Decimal places | `1` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_EnrichmentHeatmap()](../02_gene_expression/06_enrichmentheatmap.qmd) — Gene set enrichment
- [do_PathwayActivityHeatmap()](02_pathwayactivityheatmap.qmd) — Pathway activities
