# do_PathwayActivityHeatmap() | Pathway activity scores

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
sample <- sample[, sample(colnames(sample), 1000)]
```

Pathway activity heatmaps visualize **PROGENy pathway activity scores** from `decoupleR`. Shows how different signaling pathways are activated across cell populations.

---

## Basic usage

First compute pathway activities using `decoupleR`:

```{r}
#| eval: true
# Run decoupleR with PROGENy prior knowledge
net <- decoupleR::get_progeny(organism = "human", top = 500)
activities <- decoupleR::run_wmean(mat = as.matrix(sample[["RNA"]]$data), 
                                   net = net)

p <- SCpubr::do_PathwayActivityHeatmap(sample = sample,
                                       activities = activities)
p
```

---

## Group by custom variable

```{r}
#| eval: true
p <- SCpubr::do_PathwayActivityHeatmap(sample = sample,
                                       activities = activities,
                                       group.by = "subtype")
p
```

---

## Split by condition

```{r}
#| eval: true
p <- SCpubr::do_PathwayActivityHeatmap(sample = sample,
                                       activities = activities,
                                       split.by = "subtype")
p
```

---

## Show values

```{r}
#| eval: true
p <- SCpubr::do_PathwayActivityHeatmap(sample = sample,
                                       activities = activities,
                                       values.show = TRUE,
                                       values.threshold = 2)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `activities` | DecoupleR output with PROGENy | — |
| `statistic` | Activity statistic | `"norm_wmean"` |
| `return_object` | Return data along with plot | `FALSE` |

### Values display

| Parameter | Description | Default |
|-----------|-------------|---------|
| `values.show` | Show numeric values | `FALSE` |
| `values.size` | Value text size | `3` |
| `values.threshold` | Threshold on which the text changes color | `NULL` |
| `values.round` | Decimal places | `1` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_TFActivityHeatmap()](03_tfactivityheatmap.qmd) — TF activities
- [do_ActivityHeatmap()](01_activityheatmap.qmd) — Gene set activities
