# do_TFActivityHeatmap() | Transcription factor activity

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
sample <- sample[, sample(colnames(sample), 1000)]
```

TF activity heatmaps visualize **DoRothEA transcription factor activity scores** from `decoupleR`. Shows which transcription factors are most active in different cell populations.

Due to dependency problems with OmnipathR, this page will not show plots as it can not compute.
---

## Basic usage

First compute TF activities using `decoupleR`:

```{r}
#| eval: false
# Run decoupleR with DoRothEA prior knowledge
net <- decoupleR::get_dorothea(organism = "human", levels = c("A", "B", "C"))
activities <- decoupleR::run_wmean(mat = as.matrix(sample[["RNA"]]$data), 
                                   net = net)

p <- SCpubr::do_TFActivityHeatmap(sample = sample,
                                  activities = activities)
p
```

---

## Limit number of TFs

```{r}
#| eval: false
p <- SCpubr::do_TFActivityHeatmap(sample = sample,
                                  activities = activities,
                                  n_tfs = 10)
p
```

---

## Select specific TFs

```{r}
#| eval: false
p <- SCpubr::do_TFActivityHeatmap(sample = sample,
                                  activities = activities,
                                  tfs.use = c("PAX5", "STAT3", "NF-kB"))
p
```

---

## Group by custom variable

```{r}
#| eval: false
p <- SCpubr::do_TFActivityHeatmap(sample = sample,
                                  activities = activities,
                                  group.by = "cell_type")
p
```

---

## Split by condition

```{r}
#| eval: false
p <- SCpubr::do_TFActivityHeatmap(sample = sample,
                                  activities = activities,
                                  split.by = "condition")
p
```

---

## Show values

```{r}
#| eval: false
p <- SCpubr::do_TFActivityHeatmap(sample = sample,
                                  activities = activities,
                                  values.show = TRUE,
                                  values.size = 4)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `activities` | DecoupleR output with DoRothEA | — |
| `n_tfs` | Number of top TFs | `25` |
| `tfs.use` | Specific TFs to include | `NULL` |
| `statistic` | Activity statistic | `"norm_wmean"` |
| `return_object` | Return data along with plot | `FALSE` |

### Values display

| Parameter | Description | Default |
|-----------|-------------|---------|
| `values.show` | Show numeric values | `FALSE` |
| `values.size` | Value text size | `3` |
| `values.threshold` | Threshold on which the text changes color | `NULL` |
| `values.round` | Decimal places | `1` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_PathwayActivityHeatmap()](02_pathwayactivityheatmap.qmd) — Pathway activities
- [do_ActivityHeatmap()](01_activityheatmap.qmd) — Gene set activities
