# do_CorrelationHeatmap() | Feature correlation matrix

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Correlation heatmaps show **pairwise Pearson correlations** between samples, groups, or features. Use them to identify similar or distinct populations based on their expression profiles.


---

## Basic usage

Correlate cell populations using highly variable genes:

```{r}
#| eval: true
p <- SCpubr::do_CorrelationHeatmap(sample = sample)
p
```

---

## Group by custom variable

```{r}
#| eval: true
p <- SCpubr::do_CorrelationHeatmap(sample = sample,
                                   group.by = "ID")
p
```

---

## Keep diagonal

```{r}
#| eval: true
p <- SCpubr::do_CorrelationHeatmap(sample = sample,
                                   remove.diagonal = FALSE)
p
```

---

## Disable clustering

```{r}
#| eval: true
p <- SCpubr::do_CorrelationHeatmap(sample = sample,
                                   cluster = FALSE)
p
```

---

## Show values

```{r}
#| eval: true
p <- SCpubr::do_CorrelationHeatmap(sample = sample,
                                   values.show = TRUE,
                                   values.threshold = 0.5)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `input_gene_list` | Genes for correlation | `NULL` (uses HVGs) |
| `mode` | Correlation mode | `"hvg"` |
| `cluster` | Hierarchical clustering | `TRUE` |
| `remove.diagonal` | Set diagonal to NA | `TRUE` |

### Values display

| Parameter | Description | Default |
|-----------|-------------|---------|
| `values.show` | Show numeric values | `FALSE` |
| `values.size` | Value text size | `3` |
| `values.threshold` | Threshold on which the text changes color | `NULL` |
| `values.round` | Decimal places | `1` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_ExpressionHeatmap()](../02_gene_expression/05_expressionheatmap.qmd) — Expression heatmaps
- [do_GroupwiseDEHeatmap()](05_groupwisedeheatmap.qmd) — DE analysis
