# do_GroupwiseDEHeatmap() | Differential expression

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")

```

Groupwise DE heatmaps visualize the results of `Seurat::FindAllMarkers()` as a **multi-panel dot plot** showing adjusted p-values, fold changes, and mean expression per cluster.

---

## Basic usage

First run `FindAllMarkers()`:

```{r}
#| eval: true
de_genes <- Seurat::FindAllMarkers(sample, only.pos = TRUE)

p <- SCpubr::do_GroupwiseDEHeatmap(sample = sample,
                                   de_genes = de_genes)
p
```

---

## Adjust number of top genes

```{r}
#| eval: true
p <- SCpubr::do_GroupwiseDEHeatmap(sample = sample,
                                   de_genes = de_genes,
                                   top_genes = 3)
p
```

---

## Custom p-value cutoff

```{r}
#| eval: true
p <- SCpubr::do_GroupwiseDEHeatmap(sample = sample,
                                   de_genes = de_genes,
                                   p.cutoff = 0.01)
p
```

---

## Adjust dot size

```{r}
#| eval: true
p <- SCpubr::do_GroupwiseDEHeatmap(sample = sample,
                                   de_genes = de_genes,
                                   dot.scale = 6)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `de_genes` | Output from `FindAllMarkers()` | — |
| `top_genes` | Top DE genes per group | `5` |
| `p.cutoff` | Adjusted p-value threshold | `0.05` |
| `dot.scale` | Dot size multiplier | `8` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_VolcanoPlot()](07_volcanoplot.qmd) — Volcano plots
- [do_DotPlot()](../02_gene_expression/04_dotplot.qmd) — Dot plots
