# do_TermEnrichmentPlot() | GO/pathway term enrichment

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")

```

This function visualizes **enriched GO terms or pathways** from over-representation analysis (e.g., using `clusterProfiler`). Displays terms as a dot plot with gene count and p-values.

---

## Basic usage

First run enrichment analysis:

```{r}
#| eval: true
#| echo: false
# Using clusterProfiler
library("org.Hs.eg.db")
enrichment_result <- clusterProfiler::enrichGO(
  gene = Seurat::VariableFeatures(sample)[1:100],
  OrgDb = org.Hs.eg.db,
  keyType = "SYMBOL",
  ont = "BP"
)

p <- SCpubr::do_TermEnrichmentPlot(mat = enrichment_result@result)
p
```

---

## Adjust number of terms

```{r}
#| eval: true
p <- SCpubr::do_TermEnrichmentPlot(mat = enrichment_result@result,
                                   n.terms = 15)
p
```

---

## Control term wrapping

```{r}
#| eval: true
p <- SCpubr::do_TermEnrichmentPlot(mat = enrichment_result@result,
                                   n.chars = 30)  # Shorter lines
p
```

---

## Adjust dot size

```{r}
#| eval: true
p <- SCpubr::do_TermEnrichmentPlot(mat = enrichment_result@result,
                                   dot.scale = 6)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `mat` | Enrichment result (via `@result`) | — |
| `n.terms` | Number of terms to show | `25` |
| `n.chars` | Characters before wrapping | `40` |
| `dot.scale` | Dot size multiplier | `8` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_EnrichmentHeatmap()](../02_gene_expression/06_enrichmentheatmap.qmd) — Enrichment scores
- [do_GroupwiseDEHeatmap()](05_groupwisedeheatmap.qmd) — DE analysis
