# do_VolcanoPlot() | DE volcano plots

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Volcano plots display **differential expression results** showing both fold change (X-axis) and statistical significance (Y-axis). Points are colored by significance category with optional gene labels.

---

## Basic usage

First run differential expression:

```{r}
#| eval: true
de_genes <- Seurat::FindMarkers(sample, ident.1 = "IPC-like", ident.2 = "Cilia-like")

p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes)
p
```

---

## Adjust cutoffs

```{r}
#| eval: true
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes,
                            pval_cutoff = 0.01,
                            FC_cutoff = 1.5)
p
```

---

## Control gene labels

```{r}
#| eval: true
# Show more genes
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes,
                            n_genes = 10)
p
```

---

## Disable gene labels

```{r}
#| eval: true
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes,
                            add_gene_tags = FALSE)
p
```

---

## Order labels by significance or fold change

```{r}
#| eval: true
# By p-value
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes,
                            order_tags_by = "pvalue")
p
```

```{r}
#| eval: true
# By log fold change
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes,
                            order_tags_by = "logfc")
p
```

---

## Use labels instead of text

```{r}
#| eval: true
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes,
                            use_labels = TRUE)
p
```

---

## Hide threshold lines

```{r}
#| eval: true
p <- SCpubr::do_VolcanoPlot(sample = sample,
                            de_genes = de_genes,
                            plot_lines = FALSE)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `de_genes` | Output from `FindMarkers()` | — |
| `pval_cutoff` | P-value threshold | `0.05` |
| `FC_cutoff` | Fold change threshold | `2` |

### Gene labeling

| Parameter | Description | Default |
|-----------|-------------|---------|
| `add_gene_tags` | Show gene labels | `TRUE` |
| `n_genes` | Number of genes to label | `5` |
| `add_tag_side` | `"both"`, `"positive"`, `"negative"` | `"both"` |
| `order_tags_by` | `"both"`, `"pvalue"`, `"logfc"` | `"both"` |
| `use_labels` | Use `geom_label()` | `FALSE` |
| `tag_size` | Label text size | `6` |

### Threshold lines

| Parameter | Description | Default |
|-----------|-------------|---------|
| `plot_lines` | Show threshold lines | `TRUE` |
| `line_color` | Line color | `"grey75"` |
| `line_size` | Line thickness | `0.5` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_GroupwiseDEHeatmap()](05_groupwisedeheatmap.qmd) — Multi-group DE
- [do_TermEnrichmentPlot()](06_termenrichmentplot.qmd) — Enrichment analysis
