# do_CNVHeatmap() | Copy number variation

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

CNV heatmaps display **chromosomal copy number variations** across cells using output from `inferCNV`. Shows CNV scores aggregated by chromosome or chromosome arm across cell populations.

Plots will be released at a later time!
---

## Basic usage

First run `inferCNV`:

```{r}
#| eval: false
# Load chromosome locations
data("human_chr_locations", package = "SCpubr")

p <- SCpubr::do_CNVHeatmap(sample = sample,
                           infercnv_object = infercnv_obj,
                           chromosome_locations = human_chr_locations)
p
```


---

## Include chromosome arms

```{r}
#| eval: false
p <- SCpubr::do_CNVHeatmap(sample = sample,
                           infercnv_object = infercnv_obj,
                           chromosome_locations = human_chr_locations,
                           include_chr_arms = TRUE)
p
```


---

## Show values

```{r}
#| eval: false
p <- SCpubr::do_CNVHeatmap(sample = sample,
                           infercnv_object = infercnv_obj,
                           chromosome_locations = human_chr_locations,
                           values.show = TRUE)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `infercnv_object` | Output from inferCNV | — |
| `chromosome_locations` | Chromosome region data | — |
| `include_chr_arms` | Show p/q arms | `FALSE` |
| `return_object` | Return data along with plot | `FALSE` |

### Metacells

| Parameter | Description | Default |
|-----------|-------------|---------|
| `using_metacells` | inferCNV used metacells | `FALSE` |
| `metacell_mapping` | Cell-metacell mapping | `NULL` |

### Values display

| Parameter | Description | Default |
|-----------|-------------|---------|
| `values.show` | Show numeric values | `FALSE` |
| `values.size` | Value text size | `3` |
| `values.threshold` | Only show above threshold | `NULL` |
| `values.round` | Decimal places | `1` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_ExpressionHeatmap()](../02_gene_expression/05_expressionheatmap.qmd) — Expression heatmaps
