# do_LigandReceptorPlot() | Cell-cell communication

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
sample <- sample[, sample(colnames(sample), 1000)]
library(magrittr)
```

This function visualizes **ligand-receptor interactions** between cell types using output from `liana`. Shows interaction magnitude and specificity as a dot plot.

Plots will be released at a later time, when Liana's incompatibilities with Seurat v5 are solved: https://github.com/saezlab/liana/issues/200
---

## Basic usage

First run `liana`:

```{r}
#| eval: false
liana_output <- liana::liana_wrap(sample) %>% 
                liana::liana_aggregate()

p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output)
p
```

---

## Adjust number of interactions

```{r}
#| eval: false
p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output,
                                   top_interactions = 50)
p
```

---

## Sorting options

```{r}
#| eval: false
# Sort by specificity
p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output,
                                   sort.by = "A")

# Sort by magnitude
p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output,
                                   sort.by = "B")

# Sort by specificity then magnitude
p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output,
                                   sort.by = "C")

# Sort by magnitude then specificity
p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output,
                                   sort.by = "D")

# Equal weight (default)
p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output,
                                   sort.by = "E")
```

---

## Facet by ligand or receptor

```{r}
#| eval: false
# Facet by ligand
p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output,
                                   split.by = "ligand.complex")

# Facet by receptor
p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output,
                                   split.by = "receptor.complex")
```

---

## Generate chord diagrams

```{r}
#| eval: false
p <- SCpubr::do_LigandReceptorPlot(liana_output = liana_output,
                                   compute_ChordDiagrams = TRUE)
p$DotPlot
p$ChordDiagram
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `liana_output` | Aggregated liana output | — |
| `top_interactions` | Number of interactions | `25` |
| `top_interactions_by_group` | Apply per source group | `FALSE` |

### Filtering

| Parameter | Description | Default |
|-----------|-------------|---------|
| `keep_source` | Source cell types to keep | `NULL` |
| `keep_target` | Target cell types to keep | `NULL` |

### Sorting

| Parameter | Description | Default |
|-----------|-------------|---------|
| `sort.by` | Sort method (A-E) | `"E"` |
| `magnitude` | Magnitude column | `"sca.LRscore"` |
| `specificity` | Specificity column | `"aggregate_rank"` |
| `invert_specificity` | -log10 transform | `TRUE` |
| `invert_magnitude` | -log10 transform | `FALSE` |

### Appearance

| Parameter | Description | Default |
|-----------|-------------|---------|
| `dot_border` | Add dot borders | `TRUE` |
| `dot.size` | Dot size | `1` |
| `compute_ChordDiagrams` | Generate chord plots | `FALSE` |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_ChordDiagramPlot()](../03_proportions/04_chorddiagramplot.qmd) — Chord diagrams
