# do_LoadingsHeatmap() | PCA loadings

```{r}
#| results: "asis"
#| echo: false
source("../_common.R")
```

Loadings heatmaps show **which genes contribute most** to principal components. Displays top and bottom scoring genes for each PC alongside their expression levels.

---

## Basic usage

```{r}
#| eval: true
p <- SCpubr::do_LoadingsHeatmap(sample = sample)
p
```

---

## Select specific PCs

```{r}
#| eval: true
p <- SCpubr::do_LoadingsHeatmap(sample = sample,
                                dims = 1:2)
p
```

---

## Adjust number of genes

```{r}
#| eval: true
p <- SCpubr::do_LoadingsHeatmap(sample = sample,
                                top_loadings = 2)
p
```

---

## Group by cell type

```{r}
#| eval: true
p <- SCpubr::do_LoadingsHeatmap(sample = sample,
                                group.by = "subtype")
p
```

---

## Subsample cells

```{r}
#| eval: true
p <- SCpubr::do_LoadingsHeatmap(sample = sample,
                                subsample = 1000)
p
```

---

## Adjust color scales

```{r}
#| eval: true
p <- SCpubr::do_LoadingsHeatmap(sample = sample,
                                min.cutoff.loadings = -0.1,
                                max.cutoff.loadings = 0.1,
                                min.cutoff.expression = 0,
                                max.cutoff.expression = 3)
p
```

---

## Parameter reference

::: {.callout-note}
For parameters shared across many functions (color palettes, typography, legend styling, grid), see [Shared features](../00_getting_started/04_shared_features.qmd).
:::

### Core parameters

| Parameter | Description | Default |
|-----------|-------------|---------|
| `dims` | PCs to include | `1:10` |
| `top_loadings` | Top/bottom genes per PC | `5` |
| `subsample` | Number of cells to sample | `NA` |

### Scale limits

| Parameter | Description | Default |
|-----------|-------------|---------|
| `min.cutoff.loadings` | Min loading score | `NA` (q=0.05) |
| `max.cutoff.loadings` | Max loading score | `NA` (q=0.95) |
| `min.cutoff.expression` | Min expression | `NA` (0) |
| `max.cutoff.expression` | Max expression | `NA` (q=0.95) |

---

## See also

- [Shared features](../00_getting_started/04_shared_features.qmd) — Common parameters
- [do_CorrelationHeatmap()](04_correlationheatmap.qmd) — Correlation analysis
- [do_ExpressionHeatmap()](../02_gene_expression/05_expressionheatmap.qmd) — Expression heatmaps
